---
title: "IPM"
output: html_document
---

<https://github.com/ropensci/prism>
```{r}
# IPM
library(lattice)
library(fields)
library(nlme)
library(IPMpack)

# Visualization
library(ggplot2)

# Climate data
library(prism)

# Mapping
library(rgdal)
library(raster)
```

Datasets
```{r}

load("Q:/Research/Projects/SeedHarvestModeling_2019/SeedHarvestModeling/Seed_Harvest_Modeling/sim1.Rda")


```

Integral Project Models   
Merow et al. 2014 Moethods in Ecology and Evolution   

    1. Asymototic population growth rate
    2. stable stage structure
    3. reproductive values
    4. sensitivites and elasticites to matrix elements    
    
Local data, skip after loading datasets
```{r}

# A little clunky but missing code to make original sim1
sim2 <- sim1
rm(sim1)

currentyr <- as.numeric(format(as.Date(Sys.Date(),format="%Y-%m-%d"), "%Y"))
rawdatapath <- paste("Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/Asmi_Excel/Yearly Summaries/", 
                     currentyr-1,"_asmi/RawData_",
                     currentyr-1, ".csv", collapse = '', sep = '')
asmi.raw <- read.csv(path.expand(rawdatapath), na.strings = "na")
# Need to add NA to length in years before something was measured so only growth in sizeNext and treated as a new recruit, fill in NA rows for year before something was a seedling
```



# Turn AsMi data into something approrpriate for IPM
```{r}
# Need NA when dead for length
# Need NA when vegetative or seedling, dormant or dead for fruit
asmi.raw$length[asmi.raw$status %in% c("dead","dormant")] <- NA
asmi.raw$length[asmi.raw$length>100 & !is.na(asmi.raw$length)] <- NA
table(asmi.raw$fruit[asmi.raw$flower == 0]) # there are 7 times flower is 0 but fruit is 1
asmi.raw$fruit[asmi.raw$flower == 0] <- NA

# Some lengths are zero instead of NA
asmi.raw[asmi.raw$length==0 & !is.na(asmi.raw$length),]  # 23 rows! What is wrong with these?
nrow(asmi.raw[!(asmi.raw$length==0 & !is.na(asmi.raw$length)),])
nrow(asmi.raw[asmi.raw$length==0 & !is.na(asmi.raw$length),] )
asmi.raw <- asmi.raw[!(asmi.raw$length==0 & !is.na(asmi.raw$length)),]


# ps.cor <- subset(merge(asmi.age, asmi.age, 
#                        by = c("AsMi_tag_id")), year.x == year.y - 1)
ps.cor <- subset(merge(asmi.raw, asmi.raw, 
                       by = c("AsMi_tag_id")), year.x == year.y - 1)


ps.cor$surv <- 1
ps.cor$surv[ps.cor$status.y == "dead"] <- 0

table(ps.cor$surv)

# Add row of NA length for year before seedling was added
ps.seedlings <- ps.cor[ps.cor$status.x == "seedling",]
ps.seedlings$length.y <- ps.seedlings$length.x
ps.seedlings$length.x <- NA
ps.seedlings$surv <- NA
ps.corSim <- rbind(ps.cor,ps.seedlings)

# remove dormant indivduals
ps.corSim <- ps.corSim[ps.corSim$status.x != "dormant",]
ps.corSim <- ps.corSim[ps.corSim$status.y != "dormant",]



# Create simulated data
# Add climate data as climate and climateNext
sim1 <- data.frame(size = ps.corSim$length.x, sizeNext = ps.corSim$length.y, 
                   surv = ps.corSim$surv,
                   fec.seed = ps.corSim$fruit.x,
                   fec.flower = ps.corSim$flower.x,
                   year = ps.corSim$year.x,
                   yearNext = ps.corSim$year.y)

# still seems correct, some died, some have been recruited
sum(table(is.na(sim1$size)))
sum(table(is.na(sim1$sizeNext)))

# sim1 is saved on the Github repo

# save(sim1,file = "Q:/Research/Projects/SeedHarvestModeling_2019/SeedHarvestModeling/Seed_Harvest_Modeling/sim1.Rda")

head(sim1)
# This way I could split by year to see variation among years, could add in plots/sites too

sim1$fec.flower[sim1$fec.flower==0] <- NA
sim2$fec.flower[sim2$fec.flower==0] <- NA
```


Errors?  
```{r}
table(sim1$sizeNext[is.na(sim1$size)]) # five had size zero before I got rid of the ones above! Now none are zero

sim1[sim1$sizeNext == 0 & !is.na(sim1$sizeNext),]
sim2[sim2$sizeNext == 0 & !is.na(sim2$sizeNext),] # 3044 that are zero as sizeNext, should they just be NA?, probably
sim2$sizeNext[sim2$sizeNext == 0 & !is.na(sim2$sizeNext)] <- NA #Oops, should have checked if they were all surv == 0

nrow(sim1[!(is.na(sim1$size) & is.na(sim1$sizeNext)),])
sim1 <- sim1[!(is.na(sim1$size) & is.na(sim1$sizeNext)),]

```



```{r}
# load("Q:/Research/Projects/SeedHarvestModeling_2019/SeedHarvestModeling/Seed_Harvest_Modeling/sim1.Rda")

par(mfrow=c(2,2),mar=c(4,4,2,1))
plot(sim1$size,jitter(sim1$surv),xlab="Size (t)", ylab="Survival to t+1") # jittered
plot(sim1$size,sim1$sizeNext,xlab="Size (t)",ylab="Size (t+1)")
plot(sim1$size,sim1$fec.seed,xlab="Size (t)",ylab="Seed Number")
hist(sim1$sizeNext[is.na(sim1$size)],main="",xlab="Recruit Size") # probably because not true recruit
```

```{r}
par(mfrow=c(2,2),mar=c(4,4,2,1))
plot(sim2$size,jitter(sim2$surv),xlab="Size (t)", ylab="Survival to t+1") # jittered
plot(sim2$size,sim2$sizeNext,xlab="Size (t)",ylab="Size (t+1)")
plot(sim2$size,sim2$fec.seed,xlab="Size (t)",ylab="Seed Number")
hist(sim2$sizeNext[is.na(sim2$size)],main="",xlab="Recruit Size") # probably because not true recruit
```

Create dataframe of model parameters
```{r}
params=data.frame(
  surv.int=NA, # Intercept from logistic regression of survival
  surv.slope=NA, # Slope from logistic regression of survival
  growth.int=NA, # Intercept from linear regression of growth
  growth.slope=NA, # Slope from linear regression of growth
  growth.sd=NA, # Residual sd from the linear regression of growth
  seed.int=NA, # Intercept from Poisson regression of seed number
  seed.slope=NA, # Slope from Poisson regression of seed number
  recruit.size.mean=NA, # Mean recruit size
  recruit.size.sd=NA, # Standard deviation of recruit size
  establishment.prob=NA # Probability of establishment
)

# 1. survival: logistic regression
surv.reg<-glm(surv~size,data=sim1,family=binomial())
summary(surv.reg)
params$surv.int=coefficients(surv.reg)[1]
params$surv.slope=coefficients(surv.reg)[2]

# 2. growth: linear regression
growth.reg=lm(sizeNext~size,data=sim1)
params$growth.int=coefficients(growth.reg)[1]
params$growth.slope=coefficients(growth.reg)[2]
params$growth.sd=sd(resid(growth.reg))

# 3. seeds: Poisson regression
seed.reg=glm(fec.seed~size,data=sim1,family=poisson())
params$seed.int=coefficients(seed.reg)[1]
params$seed.slope=coefficients(seed.reg)[2]

# 4. size distribution of recruits
params$recruit.size.mean=mean(sim1$sizeNext[is.na(sim1$size)]) #Oh, some of the sizeNext and size are both NA! That's silly
params$recruit.size.sd=sd(sim1$sizeNext[is.na(sim1$size)])

# 5. establishment probability
# based off of one year, I need to sum over each year which I do not have in my data.frame but could add year and sum over the previous year, will want to add climate variables too  
# this is the total individuals that have length (not dormant, not dead) divided by the total fruit
params$establishment.prob=sum(is.na(sim1$size))/sum(sim1$fec.seed,na.rm=TRUE)

```




Plot, page 11
```{r}

par(mfrow=c(2,2))
xx=seq(0,100,by=.01) # sizes at which to evaluate predictions
plot(sim1$size,jitter(sim1$surv), xlab="Size (t)",ylab="Survival to t+1") # jittered
lines(xx,predict(surv.reg,data.frame(size=xx),type="response"), col="red",lwd=3)
plot(sim1$size,sim1$sizeNext,xlab="Size (t)",ylab="Size (t+1)")
lines(xx,predict(growth.reg,data.frame(size=xx)),col="red",lwd=3)
plot(sim1$size,sim1$fec.seed,xlab="Size (t)",ylab="Seed Number (t)")
lines(xx,predict(seed.reg,
data.frame(size=xx),type="response"),col="red",lwd=3)
hist(sim1$sizeNext[is.na(sim1$size)],freq=FALSE,xlab="Recruit size",main="Size of Recruits")
lines(xx,dnorm(xx,params$recruit.size.mean,
params$recruit.size.sd), col="red",lwd=3)



```



# Functions to describe life history   
     1. Use the regressions from above to build the IPM 
```{r}

# 1. survival probability function
s.x <- function(x,params) {
  u=exp(params$surv.int+params$surv.slope*x)
  return(u/(1+u))
  }

# 2. growth function
g.yx <- function(xp,x,params) {
  dnorm(xp,mean=params$growth.int+params$growth.slope*x,sd=params$growth.sd)
}

# 3. reproduction function
f.yx <- function(xp,x,params) {
  params$establishment.prob*
  dnorm(xp,mean=params$recruit.size.mean,sd=params$recruit.size.sd)*
  exp(params$seed.int+params$seed.slope*x)
}
```


# Define the boundary points (edges of cells defining the matrix), mesh points (centers of cells defining the matrix, for midpoint rule), step size (h: widths of the cells), integration limits (min.size and max.size) span the range of sizes observed in data set plus some
```{r}
min.size=.9*min(c(sim1$size,sim1$sizeNext),na.rm=T) # 0 probably wrong, want a little smaller than minimum
max.size=1.1*max(c(sim1$size,sim1$sizeNext),na.rm=T) # 110; get a little larger than largest size
n=100 # number of cells in the matrix
b <- min.size+c(0:n)*(max.size-min.size)/n # boundary points
y <- 0.5*(b[1:n]+b[2:(n+1)]) # mesh points, length 100
h <- y[2]-y[1] # step size

```

# outer() to evalute matrix at all pairwise combintations of vectors y and y to get kernel components for growth and fecundity  (???)   
    1. use midpoint rule estiamte the area under a curve
    2. heights of rectangles by outer function and width of rectanges is h   
    3. make K, the full matrix 100 x 100 cell discretization of the kernel    
    
```{r}
G <- h*outer(y,y,g.yx,params=params) # growth matrix
S <- s.x(y,params=params) # survival  
F <- h*outer(y,y,f.yx,params=params) # Previously had many individuals that were NA for both size and sizeNext, that broke this
P <- G # placeholder; redefine P on the next line
for(i in 1:n) P[,i]=G[,i]*S[i] # growth/survival matrix
K=P+F # full matrix

```

# get eigenvalues (</lambda) and eigenvectors (v-left, w-right)   
     1. lambda = asympotoic population growth rate (the dominant eigenvalue)
     2. w = stable stage distribution   
     3. v = reproductive value when normalized     

```{r}
lam <- Re(eigen(K)$values[1])
w.eigen <- Re(eigen(K)$vectors[,1])
stable.dist <- w.eigen/sum(w.eigen)
v.eigen <- Re(eigen(t(K))$vectors[,1])
repro.val <- v.eigen/v.eigen[1]

# Use eigen values to get sensitiviy and elasticity
v.dot.w=sum(stable.dist*repro.val)*h
sens=outer(repro.val,stable.dist)/v.dot.w
elas=matrix(as.vector(sens)*as.vector(K)/lam,nrow=n)
```

```{r}
par(mfrow=c(2,3),mar=c(4,5,2,2))
image.plot(y,y,t(K), xlab="Size (t)",ylab="Size (t+1)",
col=topo.colors(100), main="IPM matrix")
contour(y,y,t(K), add = TRUE, drawlabels = TRUE)
plot(y,stable.dist,xlab="Size",type="l",main="Stable size distribution")
plot(y,repro.val,xlab="Size",type="l",main="Reproductive values")
image.plot(y,y,t(elas),xlab="Size (t)",ylab="Size (t+1)",main="Elasticity")
image.plot(y,y,t(sens),xlab="Size (t)",ylab="Size (t+1)", main="Sensitivity")

```

# Improvments   
     1. add variance in growth among individuals a function of size   
     2. quadratic term for size in growth function to capture plateauing pattern in growth data (to allow for plateauing pattern)    
     3. include flowering probability in fecundity - not all individuals reproduce
```{r}
plot(growth.reg$model$size,abs(resid(growth.reg)),xlab="size",ylab="residual")

```

```{r}
 # build a regression on the residuals of growth regression as function of size using GLS (easiest)
growth.reg=gls(sizeNext~size,weights=varExp(),na.action=na.omit, data=sim1)
summary(growth.reg)
plot(sim1$size,sim1$sizeNext,main="Growth/Shrinkage/Stasis")
lines(xx,predict(growth.reg,data.frame(size=xx)),col="red",lwd=3)
```

modify params data frame, coefficients called growth.sd.int and growth.sd.slope 
```{r}
params$growth.int=coefficients(growth.reg)[1]
params$growth.slope=coefficients(growth.reg)[2]
params$growth.sigma2=summary(growth.reg)$sigma^2
params$growth.sigma2.exp=as.numeric(growth.reg$modelStruct$varStruct)

```
Let the growth function g.xy() allow standard deviation to be a function of size.  
```{r}
g.yx <- function(xp,x,params) {
  dnorm(xp,mean=params$growth.int+params$growth.slope*x,
  sd=sqrt(params$growth.sigma2*exp(2*params$growth.sigma2.exp*x)))
}

```

Rerun code to get lambda, sensitivies, elasticities
```{r}
G=h*outer(y,y,g.yx,params=params) # growth matrix
S=s.x(y,params=params) # survival
P=G # placeholder; redefine P on the next line
for(i in 1:n) P[,i]=G[,i]*S[i] # growth/survival matrix
F=h*outer(y,y,f.yx,params=params) # reproduction matrix
K=P+F # full matrix
(lam=Re(eigen(K)$values[1])) # new population growth rate
w.eigen=Re(eigen(K)$vectors[,1])
stable.dist=w.eigen/sum(w.eigen)
v.eigen=Re(eigen(t(K))$vectors[,1])
repro.val=v.eigen/v.eigen[1]
v.dot.w=sum(stable.dist*repro.val)*h
sens=outer(repro.val,stable.dist)/v.dot.w
elas=matrix(as.vector(sens)*as.vector(K)/lam,nrow=n)

```

```{r}
par(mfrow=c(2,3),mar=c(4,5,2,2))
image.plot(y,y,t(K), xlab="Size (t)",ylab="Size (t+1)",
col=topo.colors(100), main="IPM matrix")
contour(y,y,t(K), add = TRUE, drawlabels = TRUE)
plot(y,stable.dist,xlab="Size",type="l",main="Stable size distribution")
plot(y,repro.val,xlab="Size",type="l",main="Reproductive values")
image.plot(y,y,t(elas),xlab="Size (t)",ylab="Size (t+1)",main="Elasticity")
image.plot(y,y,t(sens),xlab="Size (t)",ylab="Size (t+1)", main="Sensitivity")

```

   
Check that integration happens over all size classes possible. Avoid 'eviction' where sizes are missed because are beyond the thesholds choosen. 
```{r}
plot(y,s.x(y,params),xlab="Size",type="l",
ylab="Survival Probability",lwd=12)
points(y,apply(P,2,sum),col="red",lwd=3,cex=.1,pch=19)

```

Oops, yes, column sums are lower than suvival model suggests.

See Williams et al. 2012 for ways to deal with eviction. Can add the larger individuals back into the largest stage, smaller into the smallest. 
Loop to correct for large and for offspring
```{r}
G=h*outer(y,y,g.yx,params=params) # growth matrix
S=s.x(y,params=params)
P=G
# fix eviction of offspring
for(i in 1:(n/2)) {
G[1,i]<-G[1,i]+1-sum(G[,i])
P[,i]<-G[,i]*S[i]
}
# fix eviction of large adults
for(i in (n/2+1):n) {
G[n,i]<-G[n,i]+1-sum(G[,i])
P[,i]<-G[,i]*S[i]
}
F=h*outer(y,y,f.yx,params=params) # reproduction matrix
K=P+F # full matrix
(lam=Re(eigen(K)$values[1])) # new population growth rate


```


Wow, it worked!
```{r}
plot(y,s.x(y,params),xlab="Size",type="l",
ylab="Survival Probability",lwd=12)
points(y,apply(P,2,sum),col="red",lwd=3,cex=.1,pch=19)

```

Allow for a slight downturn in growth by adding a quadratic term "I(size^2)"   
```{r}
growth.reg=gls(sizeNext~size+I(size^2),weights=varExp(), na.action=na.omit, data=sim1)
summary(growth.reg)

```

Add it into the params object
```{r}
params$growth.int=coefficients(growth.reg)[1]
params$growth.slope=coefficients(growth.reg)[2]
params$growth.sqrd=coefficients(growth.reg)[3]
params$growth.sigma2=summary(growth.reg)$sigma^2
params$growth.sigma2.exp=as.numeric(growth.reg$modelStruct$varStruct)
g.yx=function(xp,x,params) {
dnorm(xp,
mean=params$growth.int+params$growth.slope*x+params$growth.sqrd*x^2,
sd=sqrt(params$growth.sigma2*exp(2*params$growth.sigma2.exp*x)))
}

```

We rerun the code to build the matrix and obtain l, sensitivities, elasticities, etc.  
```{r}
G=h*outer(y,y,g.yx,params=params) # growth matrix
S=s.x(y,params=params) # survival
P=G # placeholder; redefine P on the next line
# fix eviction of offspring
for(i in 1:(n/2)) {
G[1,i]<-G[1,i]+1-sum(G[,i])
P[,i]<-G[,i]*S[i]
}
# fix eviction of large adults
for(i in (n/2+1):n) {
G[n,i]<-G[n,i]+1-sum(G[,i])
P[,i]<-G[,i]*S[i]
}
#for(i in 1:n) P[,i]=G[,i]*S[i] # growth/survival matrix
F=h*outer(y,y,f.yx,params=params) # reproduction matrix
K=P+F # full matrix
(lam=Re(eigen(K)$values[1])) # new population growth rate
```


```{r}
w.eigen=Re(eigen(K)$vectors[,1])
stable.dist=w.eigen/sum(w.eigen)
v.eigen=Re(eigen(t(K))$vectors[,1])
repro.val=v.eigen/v.eigen[1]
v.dot.w=sum(stable.dist*repro.val)*h
sens=outer(repro.val,stable.dist)/v.dot.w
elas=matrix(as.vector(sens)*as.vector(K)/lam,nrow=n)

```

```{r}

par(mfrow=c(2,3),mar=c(4,5,2,2))
image.plot(y,y,t(K), xlab="Size (t)",ylab="Size (t+1)",
col=topo.colors(100), main="IPM matrix")
contour(y,y,t(K), add = TRUE, drawlabels = TRUE)
plot(y,stable.dist,xlab="Size",type="l",main="Stable size distribution")
plot(y,repro.val,xlab="Size",type="l",main="Reproductive values")
image.plot(y,y,t(elas),xlab="Size (t)",ylab="Size (t+1)",main="Elasticity")
image.plot(y,y,t(sens),xlab="Size (t)",ylab="Size (t+1)", main="Sensitivity")
```

lots of individuals that are dormant? 

# add probability of flowering (interoparity)   
use the reproduction kernel (f.xy()) and add a flowering probability.   
(could also use the 'zero-inflated model' but have information on flowering so can split up distinct processes of flowering and seed conditional on flowering.)
# Flowerin probability model requires 5 steps    

     1. write flowering probability function. use logistic regression (simliar to survival function) 

```{r}
p.flower.x=function(x,params) {
  u=exp(params$flower.int+params$flower.slope*x)
  return(u/(1+u))
}

```


     2. modify reproduction function (f.xy) to indclude flowring probability function... multiply argument in f.xy by p.flower.x    
     
```{r}
f.yx=function(xp,x,params) {
  p.flower.x(x,params)*
    params$establishment.prob*
    dnorm(xp,mean=params$recruit.size.mean,sd=params$recruit.size.sd)*
    exp(params$seed.int+params$seed.slope*x)
}


```

    3. fit logistic regression fo rflowering probability. (in as asmi.raw$flower) get the slope and intercept

```{r}
table(asmi.raw$flower)
table(sim1$fec.flower)

flower.reg <- glm(fec.flower~size,data=sim1,family=binomial())
params$flower.int=coefficients(flower.reg)[1]
params$flower.slope=coefficients(flower.reg)[2]
```

     4. Update regression for seed number to include only individuals that flowered   
     
```{r}
seed.reg <- glm(fec.seed~size,data=sim1[sim1$fec.flower==1,],family=poisson())
params$seed.int=coefficients(seed.reg)[1]
params$seed.slope=coefficients(seed.reg)[2]

```

# run code with full model, including additional fixes for eviction and probability of flowering
```{r}
G=h*outer(y,y,g.yx,params=params) # growth matrix
S=s.x(y,params=params) # survival
P=G # placeholder; redefine P on the next line
# fix eviction of offspring
for(i in 1:(n/2)) {
G[1,i]<-G[1,i]+1-sum(G[,i])
P[,i]<-G[,i]*S[i]
}
# fix eviction of large adults
for(i in (n/2+1):n) {
G[n,i]<-G[n,i]+1-sum(G[,i])
P[,i]<-G[,i]*S[i]
}

 for(i in 1:n) P[,i]=G[,i]*S[i] # growth/survival matrix
F=h*outer(y,y,f.yx,params=params) # reproduction matrix
K=P+F # full matrix
(lam=Re(eigen(K)$values[1])) # new population growth rate

w.eigen=Re(eigen(K)$vectors[,1])
stable.dist=w.eigen/sum(w.eigen)
v.eigen=Re(eigen(t(K))$vectors[,1])
repro.val=v.eigen/v.eigen[1]
v.dot.w=sum(stable.dist*repro.val)*h
sens=outer(repro.val,stable.dist)/v.dot.w
elas=matrix(as.vector(sens)*as.vector(K)/lam,nrow=n)


```

```{r}
par(mfrow=c(2,3),mar=c(4,5,2,2))
image.plot(y,y,t(K), xlab="Size (t)",ylab="Size (t+1)",
  col=topo.colors(100), main="IPM matrix")
contour(y,y,t(K), add = TRUE, drawlabels = TRUE)
plot(y,stable.dist,xlab="Size",type="l",main="Stable size distribution")
plot(y,repro.val,xlab="Size",type="l",main="Reproductive values")
image.plot(y,y,t(elas),xlab="Size (t)",ylab="Size (t+1)",main="Elasticity")
image.plot(y,y,t(sens),xlab="Size (t)",ylab="Size (t+1)", main="Sensitivity")
plot(y,predict(flower.reg,newdata=data.frame(size=y),type="response"),
  xlab="Size (t)", ylab="Flowering probability",type="l")


````


# Life expectancy   
Convert matrix to object class IPMmatrix
```{r}
# Pmat <- new("IPMmatrix", nDiscrete = 0, nEnvClass = 0,
#             nBigMatrix = n, nrow = n, ncol = n, meshpoints = y,
#             env.index = 0, names.discrete = )
Pmat <- new("IPMmatrix")
Pmat[, ] = P
str(Pmat)

mle<- meanLifeExpect(Pmat)
```







#SpaDES 
Individual-based integral projection models     

     1. Make processes as modules that don't interact but are scheduled (growth, reproduction, death, climate, seed harvest) **I need climate and growth, reprodution, death... to interact! or maybe there's a good year and bad year module for each growth that is scheudled as well as harvest intensities and come at differing frequencies. 
```{r}    



```

Build modules <Individual-based integral projection models>    
```{r}



```



# Integrate climate via Bayesian approach with joint likelihood and prior probability distributions of paramteters   
Follow Oppel et al. 2014 Ecology
```{r}
library(R2WinBUGS)

# Set path to WinBUGS
bugs.dir <- "C:\\Program Files\\winbugs14_full_patched\\WinBUGS14\\WinBUGS14"

load("Q:/Research/Projects/SeedHarvestModeling_2019/SeedHarvestModeling/Seed_Harvest_Modeling/MontserratOriole_data.RData")

y 						## count data index of the number of Montserrat Orioles counted at 50 point count stations from 2000 - 2013
J						## number of fledglings observed in years where intensive nest monitoring was carried out
R						## number of breeding pairs in years where intensive nest monitoring was carried out
m						## m-array recapture matrix for ADULT Montserrat Orioles individually color-banded
mj						## m-array recapture matrix for JUVENILE Montserrat Orioles individually color-banded
rain						## pre-breeding rainfall in Montserrat from 2000 - 2013 in decimeters (= mm/100, to avoid numerical overflow in BUGS)
volc						## indicator variable for intense volcanic ashfall events from 2000 - 2013   

# WinBUGs code   
# diverts R output to a connection
sink("PVA.ipm.txt")
cat("
    
    model {
    
    # Priors for all parameters

# Initial populaiton sizes 
N1[1] ~ dnorm
    
    ")

```



Add climate data
```{r}
# Set wd for prism
options(prism.path = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/PRISM_asmiclimate")

# Update current year's monthly climate data

#Precipitation total, rain and snow
get_prism_monthlys(type="ppt", mon = 1:12, keepZip = FALSE, years = (currentyr-1):currentyr)
#Maximum temperature (average)
get_prism_monthlys(type="tmax", mon = 1:12, keepZip = FALSE, years = (currentyr-1):currentyr)
#Minimum temperature (average)
get_prism_monthlys(type="tmin", mon = 1:12, keepZip = FALSE, years = (currentyr-1):currentyr)



```

UTM Zone 13 from ArcGIS
```{r}
asmilatlong <- read.csv("Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/Asmi_Word/2016_asmi/UTM_asmiplots.csv")
# Convert from UTM to lat/long
asmi.ll <- spTransform(SpatialPoints(cbind(asmilatlong$Easting,asmilatlong$Northing), 
                                     proj4string = CRS("+proj=utm +zone=13 + datum=WGS84")), 
                       CRS("+proj=longlat +datum=WGS84"))
asmiLL <- data.frame(coordinates(asmi.ll),asmilatlong$Tag__,asmilatlong$Tag_Commen)

ggplot(asmiLL, aes(coords.x1, coords.x2, colour = asmilatlong.Tag_Commen))+
  geom_point()
```

# Want Site, Year, Month, ppt, tmin, tmax  
```{r}
maxs <- grep("tmax", ls_prism_data(absPath=TRUE)[,2])
mins <- grep("tmin", ls_prism_data(absPath = TRUE)[,2])
ppt <- grep("ppt", ls_prism_data(absPath = TRUE)[,2])
avgTemps.max <- lapply(maxs, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = extract(rastertemps, asmiLL[,c("coords.x1","coords.x2")]), date = ls_prism_data()[x,])
})
avgTemps.min <- lapply(mins, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = extract(rastertemps, asmiLL[,c("coords.x1","coords.x2")]), date = ls_prism_data()[x,])
})
avgPrecip <- lapply(ppt, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = extract(rastertemps, asmiLL[,c("coords.x1","coords.x2")]), date = ls_prism_data()[x,])
})
mins.avg <- do.call(rbind, avgTemps.min)
maxs.avg <- do.call(rbind, avgTemps.max)
ppt.avg <- do.call(rbind, avgPrecip)
mins.avg <- data.frame(mins.avg, asmilatlong$Tag__, asmilatlong$Tag_Commen)
maxs.avg <- data.frame(maxs.avg, asmilatlong$Tag__, asmilatlong$Tag_Commen)
ppt.avg <- data.frame(ppt.avg, asmilatlong$Tag__, asmilatlong$Tag_Commen)
asmi.climate <- cbind(mins.avg, maxs.avg[,1], ppt.avg[,1])
names(asmi.climate) <- c("tmin","date","Plot","Site","tmax","Precip")

#date is a factor
asmi.climate$date <- as.character(asmi.climate$date)
#split by _ select the YYYYMM numbers
asmi.climate$Year <- as.numeric(substr(vapply(strsplit(asmi.climate$date, '_'), 
                                   function(x) x[5], character(1)), 1,4))
asmi.climate$Month <- as.numeric(substr(vapply(strsplit(asmi.climate$date, '_'), 
                                   function(x) x[5], character(1)), 5,6))
asmi.c <- asmi.climate[,-2]

# Data is taken annually in July
# Label Prev12 as 08 through 12 of the previous year and give it the year for 01-07 (of the previous year (the next year) 
asmi.c$Prev12 <- asmi.c$Year
asmi.c$Prev12[asmi.c$Month > 7] <- asmi.c$Year[asmi.c$Month > 7]+1
head(asmi.c[with(asmi.c, order(Year,Month,Site,Plot))&asmi.c$Year==1980,],1000)

# get previous 12 month averages
asmi.season <- aggregate(.~Prev12+Plot+Site, data=asmi.c[,-c(6:7)], mean)

```








#ScGl example
```{r}
sg <- read.csv(paste("Q:/Research/All_Projects_by_Species/Sclerocactus SPECIES/Sclerocactus_glaucus/R_code_ScGl/R_tables/RawData_scgl_", "2018",".csv",
                     sep=""))
table(sg$Site,sg$Year) 
#Remove rows where Site is blank, if any
sg[which(sg$Site == ""),]
if(nrow(sg[-which(sg$Site == ""),])>0){
  sg <- sg[-which(sg$Site == ""),]
  }

# Create simulated data
sim1 <- data.frame(size = rnorm(), sizeNext = ps.cor$length.y, 
                   fec.seed = ps.cor$fruit.x,
                   fec.flower = ps.cor$flower.x)

par(mfrow=c(2,2),mar=c(4,4,2,1))
plot(sim1$size,jitter(sim1$surv),xlab="Size (t)", ylab="Survival to t+1") # jittered
plot(sim1$size,sim1$sizeNext,xlab="Size (t)",ylab="Size (t+1)")
plot(sim1$size,sim1$fec.seed,xlab="Size (t)",ylab="Seed Number")
hist(sim1$sizeNext[!is.na(sim1$size)],main="",xlab="Recruit Size") # probably because not true recruit


```

