---
title: "Stochastic Matrix Models"
author: "Michelle DePrenger-Levin"
date: "January 23, 2020"
output: html_document
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```


```{r, results='hide'}
library(raster)
library(sp)
library(gstat)
library(virtualspecies)
library(DiagrammeR)
library(diagram)
library(popdemo)
library(prism)

# Bayesian 
library(R2jags)
library(rjags)
library(coda)

# MPM
library(popbio)

```

```{r}

tm <- read.csv("Q:/Research/Projects/SeedHarvestModeling_2019/SeedHarvestModeling/Seed_Harvest_Modeling/Kaye_transitionmatrices.csv")

# strsplit(gsub("[^[:alnum:]]", " ", tm$Mx[1]), " ") # remove all non-aphanumeric characters


m1 <- gsub("\\[|\\]", "", tm$Mx[1]) 
m2 <- strsplit(m1, ";")  # remove the ";" that separates rows
(m3 <- lapply(m2, function(x) strsplit(trimws(x,"l"), " "))) # split each list element after removing leading whitespace
(m4 <- matrix(as.numeric(unlist(m3)), nrow = length(m3[[1]][[1]]), byrow = TRUE))

# Mx: Annual transition matrix
# Tmx: Annual trnasition matrix with transition probabilities, no fecundity
# Nx: Vector of observed stage structures (counts in each stage); must be at the start
```

```{r}

levels(tm$SPP)

transition_matrices <- lapply(split(tm, tm$SPP), function(m){ 
  out <- lapply(1:nrow(m), function(x){
    m1 <- gsub("\\[|\\]", "", m$Mx[x]) # remove "[" and "]" 
    m2 <- strsplit(m1, ";")  # remove the ";" that separates rows
    m3 <- lapply(m2, function(x) strsplit(trimws(x,"l"), " ")) # split each list element after removing leading whitespace
    m4 <- matrix(as.numeric(unlist(m3)), nrow = length(m3[[1]][[1]]), byrow = TRUE)
    m4
    })
  out
  })

transition_matrices[[1]]
# ARFE with 6, 4, 4 per the there populations
```

Estimate the paramters of how does climate impact the transitions?  Worry about lag effects? 
```{r}
# Need to know about the species, where it is to get the climate
sp.info <- read.csv("Q:/Research/Projects/SeedHarvestModeling_2019/SeedHarvestModeling/Seed_Harvest_Modeling/Kaye_speciesinfo.csv")

pop.info <- read.csv("Q:/Research/Projects/SeedHarvestModeling_2019/SeedHarvestModeling/Seed_Harvest_Modeling/Kaye_popdata.csv")

r <- getData("worldclim",var="bio",res=10)
r@layers

plot(r@layers[[1]])
```

```{r}
options(prism.path = "Q:/Research/All_Projects_by_Species/Astragalus SPECIES/Astragalus_microcymbus/PRISM_asmiclimate")

maxs <- grep("tmax", ls_prism_data(absPath=TRUE)[,2])
mins <- grep("tmin", ls_prism_data(absPath = TRUE)[,2])
ppt <- grep("ppt", ls_prism_data(absPath = TRUE)[,2])

asmiLL <- pop.info[pop.info$SPP == levels(pop.info$SPP)[1],c("Lat","Long")]

avgTemps.max <- lapply(maxs, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = raster::extract(rastertemps, asmiLL[,c("Long","Lat")]), date = ls_prism_data()[x,])
})
avgTemps.min <- lapply(mins, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = raster::extract(rastertemps, asmiLL[,c("Long","Lat")]), date = ls_prism_data()[x,])
})
avgPrecip <- lapply(ppt, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = raster::extract(rastertemps, asmiLL[,c("Long","Lat")]), date = ls_prism_data()[x,])
})
mins.avg <- do.call(rbind, avgTemps.min)
maxs.avg <- do.call(rbind, avgTemps.max)
ppt.avg <- do.call(rbind, avgPrecip)
mins.avg <- data.frame(mins.avg, levels(pop.info$SPP)[1], pop.info$POP[pop.info$SPP == levels(pop.info$SPP)[1]])
maxs.avg <- data.frame(maxs.avg, levels(pop.info$SPP)[1], pop.info$POP[pop.info$SPP == levels(pop.info$SPP)[1]])
ppt.avg <- data.frame(ppt.avg, levels(pop.info$SPP)[1], pop.info$POP[pop.info$SPP == levels(pop.info$SPP)[1]])
mins.avg$Year <- as.numeric(substr(vapply(strsplit(as.character(mins.avg$date), '_'), 
                                   function(x) x[5], character(1)), 1,4))
mins.avg$Month <- as.numeric(substr(vapply(strsplit(as.character(mins.avg$date), '_'), 
                                   function(x) x[5], character(1)), 5,6))
maxs.avg$Year <- as.numeric(substr(vapply(strsplit(as.character(maxs.avg$date), '_'), 
                                   function(x) x[5], character(1)), 1,4))
maxs.avg$Month <- as.numeric(substr(vapply(strsplit(as.character(maxs.avg$date), '_'), 
                                   function(x) x[5], character(1)), 5,6))
ppt.avg$Year <- as.numeric(substr(vapply(strsplit(as.character(ppt.avg$date), '_'), 
                                   function(x) x[5], character(1)), 1,4))
ppt.avg$Month <- as.numeric(substr(vapply(strsplit(as.character(ppt.avg$date), '_'), 
                                   function(x) x[5], character(1)), 5,6))

climate <- merge(mins.avg, maxs.avg, by = c("Year","Month","levels.pop.info.SPP..1.","pop.info.POP.pop.info.SPP....levels.pop.info.SPP..1.."))
climate <- merge(climate, ppt.avg, by = c("Year","Month","levels.pop.info.SPP..1.","pop.info.POP.pop.info.SPP....levels.pop.info.SPP..1.."))
names(climate) <- c(names(climate)[1:2],"SPP","POP","tmin","tminprism","tmax","tmaxprism","ppt","pptprism")
```
If the matrix transitions are correlated to each other and to climate or correlated to each other given climate... 
```{r}
# Project starting pop size with first transition matrix
transition_matrices[[1]][[1]] %*% as.numeric(unlist(strsplit(gsub("\\[|\\]", "",tm$Nx[1]), " ")))

# transition_matrices[[1]][1:6] are the first site
# transition_matrices[[1]][7:10] are the second site
# transition_matrices[[1]][11:14] are the third site

climate.annual <- aggregate(.~SPP+POP+Year, merge(climate, tm[tm$POP %in% 
                                                                tm$POP[tm$SPP == levels(tm$SPP)[1]],c("SPP","POP","YR")], 
                                    by.x = c("SPP","POP","Year") ,by.y = c("SPP","POP","YR"))[,c(1:3,5,7)], mean) 
                            
climate.annual.ppt <- aggregate(.~SPP+POP+Year, merge(climate, tm[tm$POP %in% 
                                                                tm$POP[tm$SPP == levels(tm$SPP)[1]],c("SPP","POP","YR")], 
                                    by.x = c("SPP","POP","Year") ,by.y = c("SPP","POP","YR"))[,c(1:3,9)], sum) 
climate.annual <- merge(climate.annual, climate.annual.ppt, by=c("SPP","POP","Year"))               


setClass("Climate_MPM", slots=list(species="character", pop="factor", Year_t0="numeric", MPM = "list",climate="data.frame"))

# setMethod("Climate_MPM", )

class(transition_matrices[[1]][[1]])

lambda(transition_matrices[[1]][[1]])


lambda.bt <- boot.transitions(transition_matrices[[1]], iterations = 100)

lambdas <- lapply(transition_matrices[[1]], function(x){
  lambda(x)
})

climate.mpm <- new("Climate_MPM", species = levels(pop.info$SPP)[1], 
                   pop = pop.info$POP[pop.info$SPP == levels(pop.info$SPP)[1]],
                   Year_t0 = tm$YR[tm$SPP == levels(pop.info$SPP)[1]], 
                   climate =  climate.annual, MPM = transition_matrices[[1]])

MPM.df <- data.frame(t(sapply(transition_matrices[[1]], function(x) as.vector(x))))
names(MPM.df) <- c("f11","t12","t13","t14","f21","t22","t23","t24","f31","t32","t33","t34","f41","t42","t43","t44")
lMPM <- data.frame(lambdas = unlist(lambdas), MPM.df, climate.annual)

# Correlations within the MPM
library(corrplot)
corrplot(cor(MPM.df), method="circle")



summary(lm(lambdas~f11+f21+f31+f41+tmin+tmax+ppt , data=lMPM))

# AIC among which climate variable best explains variation among transtion rates




lm.models <- lapply(unique.target, function(x) {lm(as.formula(paste(x,"~ .", sep = "")), 
      data = data.frame(valuesMatrix[,colnames(valuesMatrix) %in% 
                                       c(x,as.character(modelMatrix$source[which(modelMatrix$target==x)]))]))})

```



MatrixSplit: Divided means A broken into process-based submatrices U (survival), F (sexual reproduction), and C (clonal reproduction)  
Different conditions of the matrices by $MatrixTreatment
```{r}


load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/PVA/Paniw etal/patterns_temporal_autocorrelation-master/COMPADRE_v.4.0.0.RData")

table(compadre$metadata$SpeciesAccepted) # how many matrixes there are per species
hist(table(compadre$metadata$SpeciesAccepted))


# example from Wall et al. 2012 
compadre$mat[compadre$metadata$SpeciesAccepted == "Astragalus michauxii"]
compadre$metadata[compadre$metadata$SpeciesAccepted == "Astragalus michauxii",] # 4 for a three year study, all same location, one unmanipulated, one burned current year, 1yr post, 2yr post

```


Find datasets with multiple survival matrices   
```{r}
# Want only species that have more than one matU
matU_all <- mapply("[[", compadre$mat, 2)

```


Following Fieberg and Ellner 2001 example    
Random transition matrix (RTM)    
Parametric matrix method (PMM)    
Tuljapurkar's small fluctuations approximation (SFA)      
```{r}



# year since fire (instead of water level) x(t)
# Survival of each of the 4 size classes (using Astragalus michauxii) 
# t is year or space for time

t <- 1:4
# alpha is the intercept, the survival average, logistic response, of each stage class 
matU <- mapply("[[", compadre$mat[compadre$metadata$SpeciesAccepted == "Astragalus michauxii"], 2)
matU.mat <- rowSums(matrix(matU[,1], nrow = 4, byrow = FALSE)) # As the survival of each class???
# I think instead it is either averaged or just different group measures of survival
y <- lapply(1:4, function(x) matU.mat[x,])

# But only have survival matrices for 1
# S <- array(matrix(NA, nrow=4, ncol=1), dim = list(4, 1, length(t)))


set.seed(1234)

# unburned, current year, 1 yr post, 2 yr post, environment/habitat differences  
(x <- compadre$metadata$MatrixTreatment[compadre$metadata$SpeciesAccepted == "Astragalus michauxii"])

# # intercept
# alpha <- rbeta(1,1,1)
# beta1 <- runif(1, -1, 1)
# eps <- rbeta(1, 1, 1)

S[,t,x] <- 1/(1 + exp(-))

# JAGS models  
jagsdata1 <- with(data.frame(x1 = x, x2 = ))

for(t in 1:4){
  for(firefactor in 1:4){
    # Enter survival of the four stages per transition matrix of the four conditions 
    S[, x, t] <- (1 + exp())
  }
}


```




Average length of studies (N+1):number of transitions (N)
```{r}
# Years would be N+1
hist(compadre$metadata$StudyDuration, breaks = 100)
mean(compadre$metadata$StudyDuration, na.rm = TRUE) # 7.07

# The projection interal, time step of the transition matrix. as a proportion of a year; i.e. 0.1 would be 36.5 days; 0.1666 would be 60 days; 0.5  would be 6 month
table(compadre$metadata$AnnualPeriodicity)

table(compadre$metadata$MatrixDimension)


table(compadre$metadata$MatrixCriteriaOntogeny)


table(compadre$metadata$OrganismType)

# brief description of experimental treatments of the population intentially imposed by humans
table(compadre$metadata$MatrixTreatment)
```


Get the A matrices   
```{r}
AsMi_A <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Astragalus michauxii"], '[[', 1)


```


Pull perennials and annuals 
```{r}

table(compadre$metadata$OrganismType)
write.csv(unique(compadre$metadata[grep("harvest" ,compadre$metadata$MatrixTreatment),
                         c("SpeciesAccepted", "OrganismType", "Authors","Journal","YearPublication")]), 
          "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/PVA/COMPADRE_harvestmatrices.csv")

write.csv(perennial_com <- compadre$metadata[grepl("perennial", compadre$metadata$OrganismType) &
                    compadre$metadata$StudyDuration > 2,], 
          "C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/PVA/COMPADRE_perennialmatrices.csv")
table(perennial_com$MatrixTreatment)

perennial_com_harvest <- perennial_com[grep("harvest", perennial_com$MatrixTreatment),]
table(perennial_com_harvest$SpeciesAccepted)

perennial_com <- perennial_com[grepl("Unmanipulated",perennial_com$MatrixTreatment),]
perennial_com <- perennial_com[grep("Divided", perennial_com$MatrixSplit, invert = TRUE),] 
# don't want ones where process (survival) seperated from fertility


annual_com <- compadre$metadata[grepl("Annual", compadre$metadata$OrganismType) &
                    compadre$metadata$StudyDuration > 2,]
table(annual_com$MatrixTreatment)

annual_com <- annual_com[grepl("Unmanipulated",annual_com$MatrixTreatment),]
annual_com <- annual_com[grep("Divided", annual_com$MatrixSplit, invert = TRUE),] 

table(perennial_com$SpeciesAccepted)
table(annual_com$SpeciesAccepted)

```




