---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
rm(list=ls())
# install.packages("popbio")
library(popbio)
library(ggplot2)
# install.packages("IPMpack")
# library(IPMpack)
# install.packages("DirichletReg")
library(DirichletReg)
library(patchwork)
```


To multiply matrices 
> v1 <- c(1,2,3)
> v2 <- matrix(c(3,1,2,2,1,3,3,2,1), ncol = 3, byrow = TRUE)
> v1 %*% t(v2)   


or crossprod()    

# Either use popbio to estimate from virtual plants counts or just start with the projection matrix for good and bad years   
Then are larger numbers of seeds the 'good years' or need to make there be good fertility leading to lots of seeds    
How to add in life, diminishing survival after life length number of years. 
```{r}
# 100 years, long persistence
# seeds_produced <- rgamma(101, 9000, 5)
seeds_produced <- runif(101,0,10000)
virtualplant <- data.frame(year = 1:(length(seeds_produced)), seed = seeds_produced, seedling = 0, vegetative = 0, reproductive = 0, year_goodbad = NA, fecundity = seeds_produced)
virtualplant$year_goodbad[1] <- "G"

for(x in 2:(length(seeds_produced))){
  # More seeds = good year, good for seed production might or might not be better for seed germination
  # Some percent of seed will germinate
  # good year survival rates
  survrate_good <- runif(1,0.5,1)
  growthrate_good <- runif(1,0.5,1)
  # bad year survival and growth
  survrate_bad <- runif(1,0,0.5)
  growthrate_bad <- runif(1,0,0.5)
  # good or bad year
  year_goodbad <- sample(rep(c("G","G","B"),100),size = 1)##bernoulli needed
  virtualplant$year_goodbad[x] <- year_goodbad
  if(year_goodbad == "G"){
    survrate <- survrate_good
    growthrate <- growthrate_good
  } else {
    survrate <- survrate_bad
    growthrate <- growthrate_bad
  }
  # A small fraction of the seeds will germinate
  virtualplant$seedling[x] <- runif(1,0,0.05)*virtualplant$seed[x]
  # Other seeds will die, most will survive with Beta(0.9,0.2) 
  virtualplant$seed[x] <- ((virtualplant$seed[x]-virtualplant$seedling[x])*rbeta(1,.5,.2))+
    (virtualplant$seed[x-1]*survrate)
  # some seedlings will survive to be vegetative, Years good for germination likely better for survival to vegetative
  virtualplant$vegetative[x] <- (rbeta(1,growthrate,0.2)*virtualplant$seedling[x])+
    (virtualplant$vegetative[x-1]*survrate)
  # similar vegetative to reproductive (check on based on Demographic data), but lower survival of reproductive
  virtualplant$reproductive[x] <- (rbeta(1,growthrate, 0.2)*virtualplant$vegetative[x])+
    (virtualplant$reproductive[x-1]*(rbeta(1,survrate,0.9)))
}

# hist(rgamma(100,9000,5))
# hist(rbeta(100,.9,.2))
# hist(rbeta(100,growthrate,0.9))
# hist(rbeta(100,growthrate*0.5,0.9))
# hist(rnorm(100, 0.5,0.05))

plot(virtualplant$year, rowSums(virtualplant[,3:5]), type="l")

ggplot(virtualplant)+
  # geom_point(aes(year, reproductive, colour=year_goodbad))+
  # geom_point(aes(year, seedling, colour=year_goodbad))+
  # geom_point(aes(year, vegetative, colour=year_goodbad))+
  geom_line(aes(year, reproductive), colour="grey")+
  geom_line(aes(year, seedling), colour="blue")+
  geom_line(aes(year, vegetative), colour="green")+
  theme_bw()
```

# Sample from the virtual population
```{r}


```

# or start with individuals and make them stage-fate marked already   
```{r}
# 100 years, long persistence
# seeds_produced <- rgamma(101, 9000, 5)
seeds_produced <- runif(101,0,10000)
virtualplant_ind <- data.frame(year = 2, fate = NA, 
                               year_goodbad = NA, 
                               year.1 = 1, stage = NA)
virtualplant_ind$stage <- ordered(virtualplant_ind$stage, 
                                  levels = c("seed","seedling","vegetative","reproductive"))
virtualplant_ind$fate <- ordered(virtualplant_ind$fate, 
                                  levels = c("seed","seedling","vegetative","reproductive"))
virtualplant_ind$year_goodbad[1] <- "G"

for(x in 2:(length(seeds_produced))){
  # More seeds = good year, good for seed production might or might not be better for seed germination
  # Some percent of seed will germinate
  # good year survival rates
  survrate_good <- runif(1,0.5,1)
  growthrate_good <- runif(1,0.5,1)
  # bad year survival and growth
  survrate_bad <- runif(1,0,0.5)
  growthrate_bad <- runif(1,0,0.5)
  # good or bad year
  # good or bad probabilities
  y_p <- 0.25
  year_goodbad <- sample(c("G","B"), 1, replace = TRUE, prob=c(y_p,1-y_p)) ##bernoulli needed
  virtualplant$year_goodbad[x] <- year_goodbad
  if(year_goodbad == "G"){
    survrate <- survrate_good
    growthrate <- growthrate_good
  } else {
    survrate <- survrate_bad
    growthrate <- growthrate_bad
  }
  # A small fraction of the seeds will germinate
  for(i in 1:seeds_produced[x]){
    virtualplant_ind <- rbind(virtualplant_ind,
                              )
  }
  virtualplant_ind$stage <- virtualplant$seedling[x] <- runif(1,0,0.05)*virtualplant$seed[x]
  # Other seeds will die, most will survive with Beta(0.9,0.2) 
  virtualplant$seed[x] <- ((virtualplant$seed[x]-virtualplant$seedling[x])*rbeta(1,.5,.2))+
    (virtualplant$seed[x-1]*survrate)
  # some seedlings will survive to be vegetative, Years good for germination likely better for survival to vegetative
  virtualplant$vegetative[x] <- (rbeta(1,growthrate,0.2)*virtualplant$seedling[x])+
    (virtualplant$vegetative[x-1]*survrate)
  # similar vegetative to reproductive (check on based on Demographic data), but lower survival of reproductive
  virtualplant$reproductive[x] <- (rbeta(1,growthrate, 0.2)*virtualplant$vegetative[x])+
    (virtualplant$reproductive[x-1]*(rbeta(1,survrate,0.9)))
}

# hist(rbinom(1:2,100,0.6 ))
# hist(rgamma(100,9000,5))
# hist(rbeta(100,.9,.2))
# hist(rbeta(100,growthrate,0.9))
# hist(rbeta(100,growthrate*0.5,0.9))
# hist(rnorm(100, 0.5,0.05))

plot(virtualplant$year, rowSums(virtualplant[,3:5]), type="l")

ggplot(virtualplant)+
  # geom_point(aes(year, reproductive, colour=year_goodbad))+
  # geom_point(aes(year, seedling, colour=year_goodbad))+
  # geom_point(aes(year, vegetative, colour=year_goodbad))+
  geom_line(aes(year, reproductive), colour="grey")+
  geom_line(aes(year, seedling), colour="blue")+
  geom_line(aes(year, vegetative), colour="green")+
  theme_bw()

```

Projection matrix wants each row is an individual. I have everything in percents.    
Is this just percent stasis 
```{r}

# make each year the number of indivdiuals in each category that stay that transition...
# seeds_produced is year 1 through 101 but is added from year 2?
virtualplant_trans <- data.frame(virtualplant[-1,],virtualplant[-nrow(virtualplant),])

# Need to round to whole number?
# seed, seedling, vegetative, reproductive   

#percent seed to seed; ignore the reproductive to seed
A_matrix <- matrix(c(mean(((virtualplant_trans$seed-virtualplant_trans$fecundity) - virtualplant_trans$seed.1)/virtualplant_trans$seed.1), 
    #percent seed to seedling
                     mean((virtualplant_trans$seedling - virtualplant_trans$seed.1)/virtualplant_trans$seed.1), 
    #percent seed to vegetative
                     mean((virtualplant_trans$vegetative - virtualplant_trans$seed.1)/virtualplant_trans$seed.1),
    #percent seed to reproductive
                     mean((virtualplant_trans$reproductive - virtualplant_trans$seed.1)/virtualplant_trans$seed.1),
    #percent seedling to seed
                     0, 
    #seedling to vegetative
                     0, 
    #percent seedling to vegetative
                     mean((virtualplant_trans$vegetative-virtualplant_trans$seedling.1)/virtualplant_trans$seedling.1), 
    #percent seedling to reproductive
                     mean((virtualplant_trans$reproductive - virtualplant_trans$seedling.1)/virtualplant_trans$seedling.1), 
    #percent vegetative to seed
                     0,
    #percent vegetative to seedling
                     0,
    #percent vegetative to vegetative
                     mean((virtualplant_trans$vegetative - virtualplant_trans$vegetative.1)/virtualplant_trans$vegetative.1),
    #percent vegetative to reproductive
                     mean((virtualplant_trans$reproductive - virtualplant_trans$vegetative.1)/virtualplant_trans$vegetative.1),
    #percent reproductive to seed; poop, is this where or do I ignore the fecundity in this matrix? What now! It's not the number, it's the bigger number of importance somehow...
                     mean((virtualplant_trans$fecundity.1/virtualplant_trans$reproductive.1)*(virtualplant_trans$seed-virtualplant_trans$fecundity)),
    #percent reproductive to seedling
                     0,
    #percent reproductive to vegetative
                     0,
    #percent reprodutive to reproductive
                     mean((virtualplant_trans$reproductive - virtualplant_trans$reproductive.1)/virtualplant_trans$reproductive.1)), 4,4) 

  
A_matrix
```

```{r}

eigen.analysis()
```





General Virtual plant types with long seed bank persistence to short to no seed bank   
What are the differences in a good year and bad and in time since event (fire, introduction, establishment, disturbance)
```{r}
# Either directional or with retrogression
stages <- c("seed","seedling","veg","rep")

A_pers_good <- matrix(c())



```


# Integral Projection Model     
<https://cmerow.github.io/RDataScience/21_assets/Intro_to_IPMs_short.pdf>   

     1. Plot Growth as some measure of size   
     2. Plot probability of survival over sizes   
     3. Plot flowering probability, log(seed #), and recruits over size  
     4. Vital rate regressions (Growth over sizes as $\beta_0 + \beta_1 size + \beta_2 size^2$)

$$ logit[s(x)] = log\left(\frac{s(x)}{1-s(x)}\right) = \beta_1 + \beta_2 x$$

glm(y ~ x, family=binomial) to get $\beta$ and then write the s(x) function

$$s(x) = \frac{exp(\beta_1 + \beta_2 x)}{1 + exp(\beta_1 + \beta_2 x)}$$
Exponential variance function gls(s1~s0, weight=varExp(form=~fitted(.)))
```{r}

```


## Growth function g(x,y)   
g(x,y) = dnorm($\mu_y, \sigma_y$)
$\mu = \beta_1 + \beta_2 x$
$\sigma_y = f(x, covariates, ...)$

lm(y~x) to get $\beta s~and~\sigma_y$

## Fertility function f(x,y)   
glm(y~x, binomial)   
glm(y~x, Poisson) to get $\beta s$ and $\mu_{sdl}$ and $\sigma_{sdl}$ from data; dnorm for f(x,y)    

e <- eigen(M) # eigenanalysis
lambda <- Re(e$values[1]) # dominant eigenvalues   

### right eigenvector
w<-Re(e$vectors[,1]) # stable (st)age distribution
w<-v/sum(v) # standardize to total density   

### left eigenvector
et<- eigen(t(M)
v<- Re(et$vectors[,1])
v<-w/w[1] # reproductive value

# -------------------------------------------------------------------------------------------
# 2020-12-22    
# randomly generated MPM, then determine generation span  

```{r}
# lognormal for reproduction (makes more sense the Poisson, it's not a count)
# Direichlet for transitions: 2 or more elements continuous random variables [0,1] and that sum to 1; need to keep column sums less than 1 since some die
# for rare plants, constrain f and survival low but keep tradeoff like in Takada and Kawai 2020
# first way that doesn't constrain f f <- rlnorm(1 ) # Poisson lambda: vector of (non-negative) means

# Testing curves of fecundity and adult survival (s) 
# s <- seq(0,0.8,length.out = 100)
# f <- -12.5 * ((s + 0.1)^2) + 10.125 # This looks like Salguero-Gomez 2017 figure 2f for critically, endangered, vulnerable
# plot(s,f)
# f2 <- -1.25* ((s+4.6)^2) + 36.45 # Salguero-Gomez 2017 figure 2f for critically, endangered, vulnerable, straight line
# plot(s,f2)
# f3 <- 12.5 * ((s-0.9)^2) - 0.125
# plot(s, f3)
# rm(f);rm(s)

  
# should be able to set up structure with alpha
# a.mat <- cbind(1:10, 5, 10:1)
# a.mat
# rdirichlet(10, a.mat)
# rdirichlet(1, 1:4)
# rdirichlet(3, runif(4, min = 1, max = 20))

# Fast is when can progress to multiple steps larger, slow is when can only progress to next stage

Tmx_annual <- lapply(1:100, function(rep){
  # three seed stages, 1 reproductive
  # lower adult survival then reproduction is higher
  s <- runif(1,0,0.8) # adult survival
  f <- -1.25 * ((s + 4.6)^2) + 36.45 + rnorm(1, mean = 0, sd = 0.1) # add some variance around the line
  t_ij <- matrix(0, nrow = 4, ncol = 4)
  t_ij[4,1:3] <- rlnorm(1, meanlog = s, sdlog = 0.5)
  t_ij[2,1] <- t_ij[3,2] <- runif(1, min = 0, max = s)
  if(any(colSums(t_ij)>=1)){
    for(i in which(colSums(t_ij)>=1)){
      t_ij[,i] <- t_ij[,i]/ (sum(t_ij[,i]) + runif(1,0.01,0.1)) # make there be some death, 80-99% survival
      } 
  }
  t_ij[1,4] <- f
  t_ij
  })


generation.time(mean(sample(Tmx_annual,6)))

x <- lapply(1:100, function(x)(lambda(mean(sample(Tmx_annual,2))))) 
boxplot(unlist(x))
x <- lapply(1:100, function(x)(generation.time(mean(sample(Tmx_annual,2))))) 
boxplot(unlist(x))
x <- lapply(1:100, function(x)(damping.ratio(mean(sample(Tmx_annual,2))))) 
boxplot(unlist(x))

generation.time(mean(Tmx_annual))
lambda(mean(Tmx_annual))
damping.ratio(mean(Tmx_annual))
```


```{r}
Tmx_iteroslow <- lapply(1:100, function(Mx){
  s <- runif(1,0,0.8) # adult survival
  f <- -1.25 * ((s + 4.6)^2) + 36.45 + rnorm(1, mean = 0, sd = 0.1) # add some variance around the line
  t_ij <- matrix(0, nrow = 4, ncol = 4)
  t_ij[2,1] <- runif(1, min = 0.01, max = 0.99)
  t_ij[3,2] <- runif(1, min = 0.01, max = 0.99)
  t_ij[4,3] <- runif(1, min = 0.01, max = 0.99)
  # stasis or retrogression
  t_ij[2,2:3] <- t_ij[3,3:4] <- t_ij[4,4] <- s
  if(any(colSums(t_ij)>=1)){
    for(i in which(colSums(t_ij)>=1)){
      t_ij[,i] <- t_ij[,i]/(sum(t_ij[,i]))
    } 
  }
  t_ij[1,4] <- f
  t_ij
})

generation.time(mean(Tmx_iteroslow))
lambda(mean(Tmx_iteroslow))
damping.ratio(mean(Tmx_iteroslow))


Tmx_semelslow <- lapply(1:100, function(Mx){
  s <- runif(1,0,0.8) # adult survival
  f <- -1.25 * ((s + 4.6)^2) + 36.45 + rnorm(1, mean = 0, sd = 0.1) # add some variance around the line
  t_ij <- matrix(0, nrow = 4, ncol = 4)
  t_ij[2,1] <- runif(1, min = 0.01, max = 0.99)
  t_ij[3,2] <- runif(1, min = 0.01, max = 0.99)
  t_ij[4,3] <- runif(1, min = 0.01, max = 0.99)
  # stasis or retrogression
  t_ij[2,2:3] <- t_ij[3,3:4] <- s
  if(any(colSums(t_ij)>=1)){
    for(i in which(colSums(t_ij)>=1)){
      t_ij[,i] <- t_ij[,i]/(sum(t_ij[,i]))
    } 
  }
  t_ij[1,4] <- f
  t_ij
})

generation.time(mean(sample(Tmx_semelslow,6)))
lambda(mean(Tmx_semelslow))
damping.ratio(mean(Tmx_semelslow))

```

```{r}
Tmx_iterofast <- lapply(1:100, function(Mx){
  s <- runif(1,0,0.8) # adult survival
  f <- -1.25 * ((s + 4.6)^2) + 36.45 + rnorm(1, mean = 0, sd = 0.1) # add some variance around the line
  t_ij <- matrix(0, nrow = 4, ncol = 4)
  t_ij[2:4,1:3] <- runif(9, min = 0.01, max = 0.99)
  t_ij[2,3] <- 0
  t_ij[2,2] <- t_ij[3,3] <-t_ij[4,4] <- s

  if(any(colSums(t_ij)>=1)){
    for(i in which(colSums(t_ij)>=1)){
      t_ij[,i] <- t_ij[,i]/(sum(t_ij[,i]))
    } 
  }
  t_ij[1,4] <- f
  t_ij
})

generation.time(Tmx_iterofast[[1]])
generation.time(mean(sample(Tmx_iterofast,6)))
generation.time(mean(Tmx_iterofast))
lambda(mean(Tmx_iterofast))
damping.ratio(mean(Tmx_iterofast))


Tmx_semelfast <- lapply(1:100, function(Mx){
  s <- runif(1,0,0.8) # adult survival
  f <- -1.25 * ((s + 4.6)^2) + 36.45 + rnorm(1, mean = 0, sd = 0.1) # add some variance around the line
  t_ij <- matrix(0, nrow = 4, ncol = 4)
  t_ij[2:4,1:3] <- runif(9, min = 0.01, max = 0.99)
  t_ij[2,3] <- 0
  t_ij[2,2] <- t_ij[3,3] <- s

  if(any(colSums(t_ij)>=1)){
    for(i in which(colSums(t_ij)>=1)){
      t_ij[,i] <- t_ij[,i]/(sum(t_ij[,i]))
    } 
  }
  t_ij[1,4] <- f
  t_ij
})

generation.time(mean(Tmx_semelfast))
lambda(mean(Tmx_semelfast))
damping.ratio(mean(Tmx_semelfast))

```

```{r}
pathstartVirtual <-  "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output_virtual/"


# Mxs <- Tmx_annual
# hist(unlist(lapply(1:length(Mxs), function(x) Mxs[[x]][1,4])))
# Nxs_annual <- lapply(1:100, function(i){
#   c(max(1,rpois(1,runif(1, 1,500))),NA,NA,max(1,rpois(1,runif(1, 1,500))))
#   })

SimSimpleVirtual(HarvestType = "NoneVirtual", MatrixType = "Annual", Mxs = Tmx_annual,
                 Nxs = Nxs_annual, Simlength = 100, generationspan = FALSE, 
                 stablestagestart = TRUE, ps = pathstartVirtual)

# 2020-12-23♪

```
2020-12-23  
```{r}
pathstartVirtual <-  "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Sim_out_10years/"

Nxs_annual <- lapply(1:100, function(i){
  c(max(1,rpois(1,runif(1, 1,500))),NA,NA,max(1,rpois(1,runif(1, 1,500))))
  })


SimSimpleVirtual(HarvestType = "NoneVirtual50Gen", MatrixType = "Annual", Mxs = Tmx_annual,
                 Nxs = Nxs_annual, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

Nxs_othertypes <- lapply(1:100, function(i){
  rpois(4,runif(4, 1,500))
})

SimSimpleVirtual(HarvestType = "NoneVirtual50Gen", MatrixType = "SemelFast", Mxs = Tmx_semelfast,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = "NoneVirtual50Gen", MatrixType = "SemelSlow", Mxs = Tmx_semelslow,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = "NoneVirtual50Gen", MatrixType = "IteroFast", Mxs = Tmx_iterofast,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = "NoneVirtual50Gen", MatrixType = "IteroSlow", Mxs = Tmx_iteroslow,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

```

# 2020-12-25 
 after adding matrix type and simulation length to output, not saving matrices
```{r}
pathstartVirtual <-  "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Sim_outvirtemp/"

# 10 years
NoHarvestVirtual <- "Virtual10yrs"

Nxs_annual <- lapply(1:100, function(i){
  c(max(1,rpois(1,runif(1, 1,500))),NA,NA,max(1,rpois(1,runif(1, 1,500))))
  })


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "Annual", Mxs = Tmx_annual,
                 Nxs = Nxs_annual, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

Nxs_othertypes <- lapply(1:100, function(i){
  rpois(4,runif(4, 1,500))
})

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelFast", Mxs = Tmx_semelfast,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelSlow", Mxs = Tmx_semelslow,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroFast", Mxs = Tmx_iterofast,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroSlow", Mxs = Tmx_iteroslow,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

rm(NoHarvestVirtual)
## 100 years
NoHarvestVirtual <- "Virtual100yrs"


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "Annual", Mxs = Tmx_annual,
                 Nxs = Nxs_annual, Simlength = 100, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelFast", Mxs = Tmx_semelfast,
                 Nxs = Nxs_othertypes, Simlength = 100, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelSlow", Mxs = Tmx_semelslow,
                 Nxs = Nxs_othertypes, Simlength = 100, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroFast", Mxs = Tmx_iterofast,
                 Nxs = Nxs_othertypes, Simlength = 100, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroSlow", Mxs = Tmx_iteroslow,
                 Nxs = Nxs_othertypes, Simlength = 100, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

rm(NoHarvestVirtual)
## 3 generations
NoHarvestVirtual <- "Virtual3gen"


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "Annual", Mxs = Tmx_annual,
                 Nxs = Nxs_annual, Simlength = 3, generationspan = TRUE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelFast", Mxs = Tmx_semelfast,
                 Nxs = Nxs_othertypes, Simlength = 3, generationspan = TRUE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelSlow", Mxs = Tmx_semelslow,
                 Nxs = Nxs_othertypes, Simlength = 3, generationspan = TRUE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroFast", Mxs = Tmx_iterofast,
                 Nxs = Nxs_othertypes, Simlength = 3, generationspan = TRUE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroSlow", Mxs = Tmx_iteroslow,
                 Nxs = Nxs_othertypes, Simlength = 3, generationspan = TRUE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

```



# add some more semelparous
```{r}
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/PVA/Paniw etal/patterns_temporal_autocorrelation-master/COMPADRE_v.4.0.0.RData")

names(compadre[[1]])
table(compadre[[1]]$OrganismType)
# Where there is no surivval of reproductive individuals


compadre[[1]][grep("Succulent", compadre[[1]]$OrganismType),]
# semelparous <- 
compadre[[2]][compadre[[1]]$SpeciesAccepted %in% c("Agave marmorata")]
compadre[[3]][compadre[[1]]$SpeciesAccepted %in% c("Agave marmorata")][[1]]$matA
compadre[[3]][compadre[[1]]$SpeciesAccepted %in% c("Agave marmorata")][[1]]$matF
compadre[[3]][compadre[[1]]$SpeciesAccepted %in% c("Agave marmorata")][[1]]$matU
dim(compadre[[3]][compadre[[1]]$SpeciesAccepted %in% c("Agave marmorata")][[1]]$matU)

#test
x <- compadre[[3]][[1]]
semelparous <- compadre[[1]][(sapply(compadre[[3]], function(x){
  x$matU[dim(x$matA)[1],dim(x$matA)[1]] == 0
})),]

semelparous[semelparous$OrganismType != "Annual" & semelparous$StudyDuration > 5 & !is.na(semelparous$StudyDuration),]
# "Tillandsia brachycaulos" long-lived semelparous threatened species!
# "Actaea elata" rare but iteroparous Mayberry-Ella 2010
# "Anthyllis vulneraria" semel or itero depending on the environment Davison et al 2010
# 2020-12-28 emperical to match 
pathstartempirical <- "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Sim_out_10years/"
pathstartempirical3gen <- "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Sim_out_3gen/"
pathstart100 <- "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output_100years/"

compadre[[2]][compadre[[1]]$SpeciesAccepted == "Tillandsia brachycaulos"]

Nx_all_TIBR <- list(rep(1,11),rep(1,11),rep(1,11),rep(1,11)) # no seedbank, NA - unobserved seedbank
Mx_all_TIBR <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Tillandsia brachycaulos"], function(x) x$matA))
TMx_all_TIBR <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Tillandsia brachycaulos"], function(x) x$matU))

SimSimple("AnnualEmpericalNone", sppVector = "TIBR", Simlength = 3, Mxs = Mx_all_TIBR[[1]], TMxs = TMx_all_TIBR[[1]], Nxs = Nx_all_TIBR, generationspan = TRUE, ps = pathstartempirical3gen)
SimSimple("AnnualEmpericalNone", sppVector = "TIBR", Simlength = 10, Mxs = Mx_all_TIBR[[1]], TMxs = TMx_all_TIBR[[1]], Nxs = Nx_all_TIBR, generationspan = FALSE, ps = pathstartempirical)
SimSimple("AnnualEmpericalNone", sppVector = "TIBR", Simlength = 100, Mxs = Mx_all_TIBR[[1]], TMxs = TMx_all_TIBR[[1]], Nxs = Nx_all_TIBR, generationspan = FALSE, ps = pathstart100)

# Doesn't seem to be semelparous either
compadre[[2]][compadre[[1]]$SpeciesAccepted == "Cleistesiopsis bifaria"]

Nx_all_CLBI <- lapply(1:17, function(x) c(NA,rep(1,11))) # no seedbank but 1st stage is dormant-ish, NA - unobserved seedbank
Mx_all_CLBI <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Cleistesiopsis bifaria"], function(x) x$matA))
TMx_all_CLBI <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Cleistesiopsis bifaria"], function(x) x$matU))


SimSimple("AnnualEmpericalNone", sppVector = "CLBI", Simlength = 3, 
          Mxs = Mx_all_CLBI[[1]], 
          Txs = TMx_all_CLBI[[1]], Nxs = Nx_all_CLBI, generationspan = TRUE, ps = pathstartempirical3gen)
SimSimple("AnnualEmpericalNone", sppVector = "CLBI", Simlength = 10, 
          Mxs = Mx_all_CLBI[[1]], 
          TMxs = TMx_all_CLBI[[1]], Nxs = Nx_all_CLBI, generationspan = FALSE, ps = pathstartempirical)
SimSimple("AnnualEmpericalNone", sppVector = "CLBI", Simlength = 100, 
          Mxs = Mx_all_CLBI[[1]], 
          TMxs = TMx_all_CLBI[[1]], Nxs = Nx_all_CLBI, generationspan = FALSE, ps = pathstart100)
```



# Add a bunch from compadre don't worry about rarity  
```{r}

head(compadre$metadata)
table(compadre$metadata$MatrixSplit)
# test
# x <- compadre$matrixClass[[100]]
# which(grepl("seed",x))
# rm(x)

CompadreSPP <- compadre$metadata[compadre$metadata$AnnualPeriodicity == 1 &
                                   compadre$metadata$MatrixTreatment == "Unmanipulated" &
                                   compadre$metadata$MatrixSplit == "Divided" &
                                   compadre$metadata$MatrixDimension,]

SPP_compadre <- names(table(CompadreSPP$SpeciesAccepted)[table(CompadreSPP$SpeciesAccepted)>4])
SPP_compadretable <- compadre$metadata[compadre$metadata$SpeciesAccepted %in% SPP_compadre,]

# find seed stages; 
taxaSeedStages <- compadre$matrixClass[compadre$metadata$SpeciesAccepted %in% SPP_compadre]

# test
# x <- taxaSeedStages[[1]]
# nr(x)
Nx_taxaCompadre <- lapply(taxaSeedStages, function(x){
  nx <- rep(1,nrow(x))
  nx[x[,1] == "prop"] <- NA
  nx
})

Mx_all_taxaCompadre <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% SPP_compadre], function(x) x$matA)
TMx_all_taxaCompadre <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% SPP_compadre], function(x) x$matU)

# Generation time for each species, lapply over each species
generationtimeCOMPADRE <- do.call(rbind,lapply(SPP_compadre[c(1:3,5:150)], function(x){
  Mx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% x], function(x) x$matA)
  TMx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% x], function(x) x$matU)

  generationtimeCOMPADRE <- mapply(function(Mx,Tx){
    Fm <- Mx - Tx
      popbio::generation.time(Mx, # r = Fm)
      r = c(unique(which(Fm > 0, arr.ind = TRUE)[,1])),
      c = c(unique(which(Fm > 0, arr.ind = TRUE)[,2])))
    }, Mx_one_taxa, TMx_one_taxa)
  data.frame(GT = generationtimeCOMPADRE, SPP = x)
  }))
# library(Rmisc)
gensummary <- summarySE(generationtimeCOMPADRE[!is.infinite(generationtimeCOMPADRE$GT),], measurevar = "GT", groupvars = "SPP", na.rm = TRUE)

longslow <- gensummary[gensummary$GT > 5,]

pathstartRandomCompadre <- "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output_COMPADRE/"
# error in simulation longslow$SPP[13:39]
sppnow <- "Actaea cordifolia"
testStages <- compadre$matrixClass[compadre$metadata$SpeciesAccepted %in% sppnow]
testA <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% sppnow], function(x) x$matA)
testU <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% sppnow], function(x) x$matU)
testNx_list <- Nx_taxaCompadre[compadre$metadata$SpeciesAccepted %in%  sppnow]
lapply(testNx_list, function(x) length(x))
lapply(testNx_list, function(x) length(x)==8)

length(testNx_list)
length(testA)
length(testU)
testStages[[1]]

# mapply(function(s,A,U) , testStages, testA, testU)

test[!sapply(test, function(x) any(is.na(x)))]
rm(test)


lapply(SPP_compadre[!(SPP_compadre %in% c("Antirrhinum subbaeticum", "Agrimonia eupatoria", "Actaea spicata"))][30:194], function(sp){
  Mx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% sp], function(x) x$matA)
  Mx_one_taxa <- Mx_one_taxa[!sapply(Mx_one_taxa, function(x) any(is.na(x)))]

  TMx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% sp], function(x) x$matU)
  TMx_one_taxa <- TMx_one_taxa[!sapply(TMx_one_taxa, function(x) any(is.na(x)))]

  Nx_one_taxa <- Nx_taxaCompadre[compadre$metadata$SpeciesAccepted %in% sp]
  
  SimSimple("COMPADREslow3gen", sppVector = sp, Simlength = 3, 
          Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = TRUE, ps = pathstartRandomCompadre)
  
  SimSimple("COMPADREslow3gen10yrs", sppVector = sp, Simlength = 10, 
            Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)
  
  SimSimple("COMPADREslow3gen100yrs", sppVector = sp, Simlength = 100, 
            Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)

})

# find this list by see for which the largest reproductive to prop is zero?

# Various matrix sizes, need to have ones that match
#test
# sp <- "Alliaria petiolata"
# stages <- 3
# rm(stages)
# test
# x <- taxaSeedStages[[1]]
# nr(x)

lapply(SPP_compadre, function(sp){
  taxaSeedStages <- compadre$matrixClass[compadre$metadata$SpeciesAccepted == sp]
  Nx_one_list <- lapply(taxaSeedStages, function(x){
    nx <- rep(1,nrow(x))
    nx[x[,1] == "prop"] <- NA
    nx
    })
  stageNums <- unlist(lapply(Nx_one_list, function(x) length(x)))
  # rle(stageNums)$values[rle(stageNums)$lengths >4] The stage numbers with at least 5 matrices of that size
  lapply(rle(stageNums)$values[rle(stageNums)$lengths >4], function(stages){
    Mx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted == sp], function(x) x$matA)[unlist(lapply(Nx_one_list, function(x) length(x) == stages))]
    Mx_one_taxa <- Mx_one_taxa[!sapply(Mx_one_taxa, function(x) any(is.na(x)))]
    TMx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted == sp], function(x) x$matU)[unlist(lapply(Nx_one_list, function(x) length(x) == stages))]
    TMx_one_taxa <- TMx_one_taxa[!sapply(Mx_one_taxa, function(x) any(is.na(x)))]
    
    SimSimple("COMPADREslow3gen", sppVector = paste(sp,stages,"stages",sep=""), Simlength = 3, 
            Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = TRUE, ps = pathstartRandomCompadre)
    
    SimSimple("COMPADREslow3gen10yrs", sppVector = paste(sp,stages,"stages",sep=""), Simlength = 10, 
              Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)
    
    SimSimple("COMPADREslow3gen100yrs", sppVector = paste(sp,stages,"stages",sep=""),Simlength = 100, 
              Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)
    })
})


# Semelparous species in COMPADRE from Takada and Kawai 2020
lapply(c("Genianella compestris","Rubus praecox","Rubus ursinus","Carlina vulgaris","Carum carvi","Digitalis purpurea","Picris hieracioides","Centaurea jacea","Cirsium dissectum","Cirsium palustre","Cirsium pitcheri","Cirsium scariosum","Cirsium tracyi","Dipsacus fullonum","Saussurea medusa","Tragapogon pratensis"), function(sp){
  Mx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% sp], function(x) x$matA)
  Mx_one_taxa <- Mx_one_taxa[!sapply(Mx_one_taxa, function(x) any(is.na(x)))]

  TMx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% sp], function(x) x$matU)
  TMx_one_taxa <- TMx_one_taxa[!sapply(TMx_one_taxa, function(x) any(is.na(x)))]

  Nx_one_taxa <- Nx_taxaCompadre[compadre$metadata$SpeciesAccepted %in% sp]
  
  SimSimple("COMPADREslow3gen", sppVector = sp, Simlength = 3, 
          Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = TRUE, ps = pathstartRandomCompadre)
  
  SimSimple("COMPADREslow3gen10yrs", sppVector = sp, Simlength = 10, 
            Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)
  
  SimSimple("COMPADREslow3gen100yrs", sppVector = sp, Simlength = 100, 
            Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)

})

```



```{r}
# Skip for now, dormant and damaged stages. 
compadre[[2]][compadre[[1]]$SpeciesAccepted == "Actaea elata"]

Nx_all_ACEL <- list(rep(1,11),rep(1,11),rep(1,11),rep(1,11)) # no seedbank, NA - unobserved seedbank
Mx_all_ACEL <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Actaea elata"], function(x) x$matA))
TMx_all_ACEL <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Tillandsia brachycaulos"], function(x) x$matU))


SimSimple("AnnualEmpericalNone", sppVector = "SEER", Simlength = 3, Mxs = Mx_all_SEER, TMxs = TMx_all_SEER, Nxs = Nx_all_SEER, generationspan = TRUE, ps = pathstartempirical3gen)
SimSimple("AnnualEmpericalNone", sppVector = "SEER", Simlength = 10, Mxs = Mx_all_SEER, TMxs = TMx_all_SEER, Nxs = Nx_all_SEER, generationspan = FALSE, ps = pathstartempirical)
SimSimple("AnnualEmpericalNone", sppVector = "SEER", Simlength = 100, Mxs = Mx_all_SEER, TMxs = TMx_all_SEER, Nxs = Nx_all_SEER, generationspan = FALSE, ps = pathstart100)

compadre[[2]][compadre[[1]]$SpeciesAccepted == "Centaurea corymbosa"] # might be semelparous but only one matrix
compadre[[2]][compadre[[1]]$SpeciesAccepted == "Furcraea parmentieri"] # none
compadre[[2]][compadre[[1]]$Genus == "Furcraea"] # none


```













