---
title: "VirtualSpecies"
author: "Michelle DePrenger-Levin"
date: "1/4/2021"
output: html_document
---


```{r}
rm(list=ls())
library(popbio)
library(ggplot2)
library(patchwork)

#automatic install of packages if they are not installed already
# list.of.packages <- c(
#   "foreach",
#   "doParallel",
#   "ranger",
#   "palmerpenguins",
#   "tidyverse",
#   "kableExtra"
#   )
# 
# new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
# 
# if(length(new.packages) > 0){
#   install.packages(new.packages, dep=TRUE)
# }

#loading packages
for(package.i in list.of.packages){
  suppressPackageStartupMessages(
    library(
      package.i, 
      character.only = TRUE
      )
    )
}

```

# -------------------------------------------------------------------------------------------
# 2020-12-22    
# randomly generated MPM, then determine generation span  
## set seed within VirtualSepeciesTypes.R

```{r}
generation.time(mean(sample(Tmx_annual,6)))

x <- lapply(1:100, function(x)(lambda(mean(sample(Tmx_annual,2))))) 
boxplot(unlist(x))
x <- lapply(1:100, function(x)(generation.time(mean(sample(Tmx_annual,2))))) 
boxplot(unlist(x))
x <- lapply(1:100, function(x)(damping.ratio(mean(sample(Tmx_annual,2))))) 
boxplot(unlist(x))

generation.time(mean(Tmx_annual))
lambda(mean(Tmx_annual))
damping.ratio(mean(Tmx_annual))

generation.time(mean(Tmx_iteroslow))
lambda(mean(Tmx_iteroslow))
damping.ratio(mean(Tmx_iteroslow))

generation.time(mean(sample(Tmx_semelslow,6)))
lambda(mean(Tmx_semelslow))
damping.ratio(mean(Tmx_semelslow))

generation.time(Tmx_iterofast[[1]])
generation.time(mean(sample(Tmx_iterofast,6)))
generation.time(mean(Tmx_iterofast))
lambda(mean(Tmx_iterofast))
damping.ratio(mean(Tmx_iterofast))


generation.time(mean(Tmx_semelfast))
lambda(mean(Tmx_semelfast))
damping.ratio(mean(Tmx_semelfast))

```

# Simulate stochastic projection with no perturbation
```{r}

library(foreach)
library(doParallel)
library(parallel)


# test
# Mx_list <- sample(Tmx_iteroslow,5)
Mx_list <- Tmx_annual
Nxstages <- c(NA,NA,NA,1)
which(lapply(Mx_list, function(x) lambda(x))<1) # which in list are growth rates < 1
## Should pick out random that have above and below 1 growth rates, not concerned if all above 1, well, maybe are with perturbations, if you're worried about extinction risk, you'd have at least some annual lambdas under 1
length(which(lapply(Tmx_annual, function(x) lambda(x))<1)) # 7
length(which(lapply(Tmx_semelfast, function(x) lambda(x))<1)) # 10
length(which(lapply(Tmx_semelslow, function(x) lambda(x))<1)) # 55
length(which(lapply(Tmx_iterofast, function(x) lambda(x))<1)) # 2
length(which(lapply(Tmx_iteroslow, function(x) lambda(x))<1)) # 31


# Nxs_annual <- lapply(1:100, function(i){
#   c(NA,NA,NA,max(1,rpois(1,runif(1, 1,500))))
#   })
StartPopSize <- 10
# vec <- rep(0,4)
MatrixType <- "Annual"
lapply(Mx_sample, function(x) lambda(x))
rm(Mx_list);rm(vec);rm(Mx_sample);rm(TMx_list); rm(StartPopSize); rm(vec); rm(Nxstages); rm(gentim); rm(projlength)

# Nx should be the stable stage distribution
# Nxstages: Only need to define seed stages in annuals, rest are not seed stages
virtualPVA <- function(Mx_list, Nxstages = rep(1,4), StartPopSize, MatrixType){
  # Take a sample of size at least 5, remove fecundity for TMx_list 
  Mx_sample <- sample(Mx_list, size = max(5,rpois(1,10)))
  TMx_list <- lapply(Mx_sample, function(i){ 
                                          i[1,4] <- 0
                                          i
                                        })
  # calculate generation time to scale by generation time or set number of years for projection length
  gentim <- popbio::generation.time(mean(Mx_sample))
  projlength <- max(gentim, 100)
  # scale the above ground growth to the starting population size
  Nx <- stable.stage(mean(Mx_sample))
  Nx_scale <- Nx[!is.na(Nxstages)]
  vec <- matrix(floor(Nx*(StartPopSize/sum(Nx_scale))), ncol = 1)
  # Initilize for loop
  out <- c()
  Extant <- c()
  yr <- c()
  mats <- list()
  popsz <- c()
  Tulapprox <- c()
  LogGrowthsim <- c()
  LogGrowthSimLowerCI <- c()
  LogGrowthSimUpperCI <- c()
  Time2Ext <- NA
  # PVA loop
  # i: 100 replicates
  n.cores <- parallel::detectCores()-1
  my.cluster <- parallel::makeCluster(
    n.cores, 
    type = "PSOCK"
    )
  #register it to be used by %dopar%
  doParallel::registerDoParallel(cl = my.cluster)
  dfout <- foreach(i = 1:100, .combine = rbind, .packages = "popbio") %dopar% {
      # Initilize for loop
        out <- c()
        Extant <- c()
        yr <- c()
        mats <- list()
        popsz <- c()
        Tulapprox <- c()
        LogGrowthsim <- c()
        LogGrowthSimLowerCI <- c()
        LogGrowthSimUpperCI <- c()
        Time2Ext <- NA
    for(i in 1:projlength){
      yr[i] <- i
      mats[[i]] <- (sample(Mx_sample,1))[[1]]
      vec <- floor(mats[[i]]%*%vec)
      popsz[i] <- sum(vec[!is.na(Nxstages)])
      Extant[i] <- if(popsz[i]<1) 0 else 1
      # A list of 2 or more, will be NA until second year
      if(i == 1){
        Tulapprox[i] <- NA
        LogGrowthsim[i] <- NA
        LogGrowthSimLowerCI[i] <- NA
        LogGrowthSimUpperCI[i] <- NA
      } else {
        stochLambda <- popbio::stoch.growth.rate(mats, verbose = FALSE)
        Tulapprox[i] <- stochLambda$approx
        LogGrowthsim[i] <-  stochLambda$sim
        LogGrowthSimLowerCI[i] <-  stochLambda$sim.CI[1]
        LogGrowthSimUpperCI[i] <-  stochLambda$sim.CI[2]
        }
      if(Extant[i] == 0) Time2Ext <- i
      if(popsz[i] == 0) break
    }
    data.frame(PopSize = popsz, Year = yr, MatType = MatrixType, GenTime = gentim,
             Extant = Extant, StPopSz = StartPopSize, Time2Extinction = Time2Ext, Tulapprox = Tulapprox,
             LogGrowthsim = LogGrowthsim, LogGrowthSimLowerCI = LogGrowthSimLowerCI, LogGrowthSimUpperCI = LogGrowthSimUpperCI,
             ProjectionLength = projlength)
  }
  parallel::stopCluster(cl = my.cluster)
  gc()
  dfout
}
```

# Just go the longer of 3 generations or 100 years to see the difference at each point.
```{r}

c(10,50,100,500)
annuals <- do.call(rbind,lapply(10, function(StPopSz){
  out <- virtualPVA(Mx_list = Tmx_annual, Nxstages = c(NA,NA,NA,1), StartPopSize = StPopSz, MatrixType = "Annual")
  save(out, file =  paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output/annuals",StPopSz,".Rdata", sep=""))
  out
  }))

iterofast <- do.call(rbind,lapply(c(10,50,100,500), function(StPopSz){
  out <- virtualPVA(Mx_list = Tmx_iterofast, StartPopSize = StPopSz, MatrixType = "IteroFast")
  save(out, file =  paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output/IteroFast",StPopSz,".Rdata", sep=""))
  out
  }))

iteroslow <- do.call(rbind,lapply(c(10,50,100,500), function(StPopSz){
  out <- virtualPVA(Mx_list = Tmx_iteroslow, StartPopSize = StPopSz, MatrixType = "IteroSlow")
  save(out, file =  paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output/IteroSlow",StPopSz,".Rdata", sep=""))
  out
  }))

semelfast <- do.call(rbind,lapply(c(10,50,100,500), function(StPopSz){
  out <- virtualPVA(Mx_list = Tmx_semelfast, StartPopSize = StPopSz, MatrixType = "semelFast")
  save(out, file =  paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output/semelFast",StPopSz,".Rdata", sep=""))
  out
  }))

semelslow <- do.call(rbind,lapply(c(10,50,100,500), function(StPopSz){
  out <- virtualPVA(Mx_list = Tmx_semelslow, StartPopSize = StPopSz, MatrixType = "semelSlow")
  save(out, file =  paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output/semelSlow",StPopSz,".Rdata", sep=""))
  out
  }))



# mapply(function(Type, Matrices){
#   Typeout <- do.call(rbind,lapply(c(10,50,100,500), function(StPopSz){
#     out <- virtualPVA(Mx_list = Matrices, StartPopSize = StPopSz, MatrixType = Type)
#     out
#     }))
#   save(Typeout, file =  paste("C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output/",Matrices,StPopSz,".Rdata",sep=""))
#   },  c("IteroFast","IteroSlow","SemelFast","SemelSlow"), list(Tmx_iterofast,Tmx_iteroslow,Tmx_semelfast,Tmx_semelslow))

```

```{r}
head(annuals)

```



## Try with simplified SeedHarvestSim function, set.seed with random generation of species types 
# errors, just do virtualPVA first, then see if simplified or just old version works. 
```{r}
pathstart <- "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output/"
Nxs_annual <- list(c(NA,NA,NA,1))
Nxs_othertypes # would be same as c(1,1,1,1)
# Test
# Mxs <- Tmx_annual
# rm(Mxs)

SimVirtualNoHarvest(ps = pathstart, MatrixType = "Annual", Mxs = Tmx_annual, Nxs = Nxs_annual, Simlength = 10, stablestagestart = TRUE, generationspan = FALSE)
SimVirtualNoHarvest(ps = pathstart, MatrixType = "Annual", Mxs = Tmx_annual, Nxs = Nxs_annual, Simlength = 100, stablestagestart = TRUE, generationspan = FALSE)
SimVirtualNoHarvest(ps = pathstart, MatrixType = "Annual", Mxs = Tmx_annual, Nxs = Nxs_annual, Simlength = 3, stablestagestart = TRUE, generationspan = TRUE)

mapply(function(TMx, MT){
  SimVirtualNoHarvest(ps = pathstart, MatrixType = MT, Mxs = TMx, Nxs = Nxs_othertypes, Simlength = 10, stablestagestart = TRUE, generationspan = FALSE)
  SimVirtualNoHarvest(ps = pathstart, MatrixType = MT, Mxs = TMx, Nxs = Nxs_othertypes, Simlength = 100, stablestagestart = TRUE, generationspan = FALSE)
  SimVirtualNoHarvest(ps = pathstart, MatrixType = MT, Mxs = TMx, Nxs = Nxs_othertypes, Simlength = 3, stablestagestart = TRUE, generationspan = TRUE)
  }, list(Tmx_iterofast,Tmx_iteroslow, Tmx_semelfast, Tmx_semelslow), c("IteroFast","IteroSlow","SemelFast","SemelSlow"))



```
  




```{r}
# sample(Tmx_annual[which(lapply(Tmx_annual, function(x) lambda(x))<1.5) ],max(5,rpois(1,10)), replace = TRUE)
alltracks <- do.call(rbind, lapply(c(10,50,100,500), function(StPopSz){
    annualtrack <- do.call(rbind, lapply(1:100, function(i){
      # virtualPVA(sample(Tmx_annual[which(lapply(Tmx_annual, function(x) lambda(x))<1.5) ],max(5,rpois(1,10)), replace = TRUE),
      virtualPVA(sample(Tmx_annual,max(5,rpois(1,10))),
                 Nxstages = c(NA,NA,NA,1),
                 StartPopSize = StPopSz, paste("Annual",i,sep="_"))
    }))
    iteroslowtrack <- do.call(rbind, lapply(1:100, function(i){
      # virtualPVA(sample(Tmx_iteroslow[which(lapply(Tmx_iteroslow, function(x) lambda(x))<1.5) ],max(5,rpois(1,10)), replace = TRUE), 
      virtualPVA(sample(Tmx_iteroslow,max(5,rpois(1,10))),
                 Nxstages = rep(1,4),
                 StartPopSize = StPopSz, paste("IteroSlow",i,sep="_"))
    }))
    iterofasttrack <- do.call(rbind, lapply(1:100, function(i){
      # virtualPVA(sample(Tmx_iterofast[which(lapply(Tmx_iterofast, function(x) lambda(x))<1.5) ],max(5,rpois(1,10)), replace = TRUE),
      virtualPVA(sample(Tmx_iterofast,max(5,rpois(1,10))),
                 Nxstages = rep(1,4),
                 StartPopSize = StPopSz, paste("IteroFast",i,sep="_"))
    }))
    semelslowtrack <- do.call(rbind, lapply(1:100, function(i){
      # virtualPVA(sample(Tmx_semelslow[which(lapply(Tmx_semelslow, function(x) lambda(x))<1.5) ],max(5,rpois(1,10)), replace = TRUE),
      virtualPVA(sample(Tmx_semelslow,max(5,rpois(1,10))),
                 Nxstages = rep(1,4),
                 StartPopSize = StPopSz, paste("SemelSlow",i,sep="_"))
      }))
    semelfasttrack <- do.call(rbind, lapply(1:100, function(i){
      # virtualPVA(sample(Tmx_semelfast[which(lapply(Tmx_semelfast, function(x) lambda(x))<1.5) ],max(5,rpois(1,10)), replace = TRUE),
      virtualPVA(sample(Tmx_semelfast,max(5,rpois(1,10))),
                 Nxstages = rep(1,4),
                 StartPopSize = StPopSz, paste("SemelFast",i,sep="_"))
      }))
    alltracks <- do.call(rbind, list(annualtrack,iteroslowtrack,iterofasttrack,semelslowtrack,semelfasttrack))
    alltracks
}))

table(alltracks$MatType, alltracks$StPopSz)
alltracks$MatrixType <- unlist(lapply(alltracks$MatType, function(x) strsplit(x, "_")[[1]][1]))
alltracks$MatrixType <- factor(alltracks$MatrixType, levels = c("Annual","SemelFast","IteroFast","SemelSlow","IteroSlow"))

# x <- split(alltracks, alltracks$MatType)[[1]]
# rm(x)
untilextirpated <- do.call(rbind,lapply(split(alltracks, alltracks$MatType), function(x){
  if(min(which(x$PopSize == 0)) <= 150){
    x[1:min(which(x$PopSize == 0)),]
    } else {
      x
      }
  }))
ggplot(untilextirpated[untilextirpated$Extant ==  0,], aes(MatrixType, Year))+ #, colour = MatrixType))+
  facet_wrap(~StPopSz)+
  geom_jitter(alpha = 0.5, aes(colour = MatrixType))+
  geom_violin(alpha = .1)+
  theme_bw()+
  coord_flip()+
  xlab("Matrix Type")+
  ylab("Time to extinction")+
  theme(legend.position = "none")
```


```{r}
# [untilextirpated$GenTime <= 5,]
ggplot(untilextirpated, aes(Year, Extant, group = MatType, colour = GenTime))+
  geom_line(alpha=0.75)+
  # scale_color_gradient2(low = "red", mid = "green", high = "blue", midpoint = 3)+
  scale_colour_continuous(type = "viridis")+
  theme_bw()+
  geom_jitter(width = 0.01, height = 0.01, alpha = 0.5, shape = 16)

ggplot(untilextirpated[untilextirpated$GenTime > 5,], aes(Year, Extant, group = MatType, colour = GenTime))+
  geom_line(alpha=0.75)+
  scale_color_gradient2(low = "blue", mid = "yellow", high = "red", midpoint = 10)+
  theme_bw()+
  geom_jitter(width = 0.01, height = 0.01, alpha = 0.5, shape = 16)
  
table(alltracks$Extant, alltracks$MatType)
```



```{r}
pathstartVirtual <-  "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output_virtual/"


SimSimpleVirtual(HarvestType = "NoneVirtual", MatrixType = "Annual", Mxs = Tmx_annual,
                 Nxs = Nxs_annual, Simlength = 100, generationspan = FALSE, 
                 stablestagestart = TRUE, ps = pathstartVirtual)

# 2020-12-23

```
2020-12-23  
```{r}
pathstartVirtual <-  "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Sim_out_10years/"

Nxs_annual <- lapply(1:100, function(i){
  # c(max(1,rpois(1,runif(1, 1,500))),NA,NA,max(1,rpois(1,runif(1, 1,500))))
  c(NA,NA,NA,max(1,rpois(1,runif(1, 1,500))))
  })


SimSimpleVirtual(HarvestType = "NoneVirtual50Gen", MatrixType = "Annual", Mxs = Tmx_annual,
                 Nxs = Nxs_annual, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

Nxs_othertypes <- lapply(1:100, function(i){
  rpois(4,runif(4, 1,500))
})

SimSimpleVirtual(HarvestType = "NoneVirtual50Gen", MatrixType = "SemelFast", Mxs = Tmx_semelfast,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = "NoneVirtual50Gen", MatrixType = "SemelSlow", Mxs = Tmx_semelslow,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = "NoneVirtual50Gen", MatrixType = "IteroFast", Mxs = Tmx_iterofast,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = "NoneVirtual50Gen", MatrixType = "IteroSlow", Mxs = Tmx_iteroslow,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

```

# 2020-12-25 
 after adding matrix type and simulation length to output, not saving matrices
```{r}
pathstartVirtual <-  "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Sim_outvirtemp/"

# 10 years
NoHarvestVirtual <- "Virtual10yrs"

Nxs_annual <- lapply(1:100, function(i){
  c(max(1,rpois(1,runif(1, 1,500))),NA,NA,max(1,rpois(1,runif(1, 1,500))))
  })


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "Annual", Mxs = Tmx_annual,
                 Nxs = Nxs_annual, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

Nxs_othertypes <- lapply(1:100, function(i){
  rpois(4,runif(4, 1,500))
})

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelFast", Mxs = Tmx_semelfast,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelSlow", Mxs = Tmx_semelslow,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroFast", Mxs = Tmx_iterofast,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroSlow", Mxs = Tmx_iteroslow,
                 Nxs = Nxs_othertypes, Simlength = 10, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

rm(NoHarvestVirtual)
## 100 years
NoHarvestVirtual <- "Virtual100yrs"


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "Annual", Mxs = Tmx_annual,
                 Nxs = Nxs_annual, Simlength = 100, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelFast", Mxs = Tmx_semelfast,
                 Nxs = Nxs_othertypes, Simlength = 100, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelSlow", Mxs = Tmx_semelslow,
                 Nxs = Nxs_othertypes, Simlength = 100, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroFast", Mxs = Tmx_iterofast,
                 Nxs = Nxs_othertypes, Simlength = 100, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroSlow", Mxs = Tmx_iteroslow,
                 Nxs = Nxs_othertypes, Simlength = 100, generationspan = FALSE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

rm(NoHarvestVirtual)
## 3 generations
NoHarvestVirtual <- "Virtual3gen"


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "Annual", Mxs = Tmx_annual,
                 Nxs = Nxs_annual, Simlength = 3, generationspan = TRUE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelFast", Mxs = Tmx_semelfast,
                 Nxs = Nxs_othertypes, Simlength = 3, generationspan = TRUE,
                 stablestagestart = TRUE, ps = pathstartVirtual)


SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "SemelSlow", Mxs = Tmx_semelslow,
                 Nxs = Nxs_othertypes, Simlength = 3, generationspan = TRUE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroFast", Mxs = Tmx_iterofast,
                 Nxs = Nxs_othertypes, Simlength = 3, generationspan = TRUE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

SimSimpleVirtual(HarvestType = NoHarvestVirtual, MatrixType = "IteroSlow", Mxs = Tmx_iteroslow,
                 Nxs = Nxs_othertypes, Simlength = 3, generationspan = TRUE,
                 stablestagestart = TRUE, ps = pathstartVirtual)

```



# add some more semelparous
```{r}
load("C:/Users/deprengm/OneDrive - Denver Botanic Gardens/P drive/My Documents/UCDenver_phd/Dissertation/PVA/Paniw etal/patterns_temporal_autocorrelation-master/COMPADRE_v.4.0.0.RData")

names(compadre[[1]])
table(compadre[[1]]$OrganismType)
# Where there is no surivval of reproductive individuals


compadre[[1]][grep("Succulent", compadre[[1]]$OrganismType),]
# semelparous <- 
compadre[[2]][compadre[[1]]$SpeciesAccepted %in% c("Agave marmorata")]
compadre[[3]][compadre[[1]]$SpeciesAccepted %in% c("Agave marmorata")][[1]]$matA
compadre[[3]][compadre[[1]]$SpeciesAccepted %in% c("Agave marmorata")][[1]]$matF
compadre[[3]][compadre[[1]]$SpeciesAccepted %in% c("Agave marmorata")][[1]]$matU
dim(compadre[[3]][compadre[[1]]$SpeciesAccepted %in% c("Agave marmorata")][[1]]$matU)

#test
x <- compadre[[3]][[1]]
semelparous <- compadre[[1]][(sapply(compadre[[3]], function(x){
  x$matU[dim(x$matA)[1],dim(x$matA)[1]] == 0
})),]

semelparous[semelparous$OrganismType != "Annual" & semelparous$StudyDuration > 5 & !is.na(semelparous$StudyDuration),]
# "Tillandsia brachycaulos" long-lived semelparous threatened species!
# "Actaea elata" rare but iteroparous Mayberry-Ella 2010
# "Anthyllis vulneraria" semel or itero depending on the environment Davison et al 2010
# 2020-12-28 emperical to match 
pathstartempirical <- "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Sim_out_10years/"
pathstartempirical3gen <- "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Sim_out_3gen/"
pathstart100 <- "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output_100years/"

compadre[[2]][compadre[[1]]$SpeciesAccepted == "Tillandsia brachycaulos"]

Nx_all_TIBR <- list(rep(1,11),rep(1,11),rep(1,11),rep(1,11)) # no seedbank, NA - unobserved seedbank
Mx_all_TIBR <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Tillandsia brachycaulos"], function(x) x$matA))
TMx_all_TIBR <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Tillandsia brachycaulos"], function(x) x$matU))

SimSimple("AnnualEmpericalNone", sppVector = "TIBR", Simlength = 3, Mxs = Mx_all_TIBR[[1]], TMxs = TMx_all_TIBR[[1]], Nxs = Nx_all_TIBR, generationspan = TRUE, ps = pathstartempirical3gen)
SimSimple("AnnualEmpericalNone", sppVector = "TIBR", Simlength = 10, Mxs = Mx_all_TIBR[[1]], TMxs = TMx_all_TIBR[[1]], Nxs = Nx_all_TIBR, generationspan = FALSE, ps = pathstartempirical)
SimSimple("AnnualEmpericalNone", sppVector = "TIBR", Simlength = 100, Mxs = Mx_all_TIBR[[1]], TMxs = TMx_all_TIBR[[1]], Nxs = Nx_all_TIBR, generationspan = FALSE, ps = pathstart100)

# Doesn't seem to be semelparous either
compadre[[2]][compadre[[1]]$SpeciesAccepted == "Cleistesiopsis bifaria"]

Nx_all_CLBI <- lapply(1:17, function(x) c(NA,rep(1,11))) # no seedbank but 1st stage is dormant-ish, NA - unobserved seedbank
Mx_all_CLBI <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Cleistesiopsis bifaria"], function(x) x$matA))
TMx_all_CLBI <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Cleistesiopsis bifaria"], function(x) x$matU))


SimSimple("AnnualEmpericalNone", sppVector = "CLBI", Simlength = 3, 
          Mxs = Mx_all_CLBI[[1]], 
          Txs = TMx_all_CLBI[[1]], Nxs = Nx_all_CLBI, generationspan = TRUE, ps = pathstartempirical3gen)
SimSimple("AnnualEmpericalNone", sppVector = "CLBI", Simlength = 10, 
          Mxs = Mx_all_CLBI[[1]], 
          TMxs = TMx_all_CLBI[[1]], Nxs = Nx_all_CLBI, generationspan = FALSE, ps = pathstartempirical)
SimSimple("AnnualEmpericalNone", sppVector = "CLBI", Simlength = 100, 
          Mxs = Mx_all_CLBI[[1]], 
          TMxs = TMx_all_CLBI[[1]], Nxs = Nx_all_CLBI, generationspan = FALSE, ps = pathstart100)
```



# Add a bunch from compadre don't worry about rarity  
```{r}

head(compadre$metadata)
table(compadre$metadata$MatrixSplit)
# test
# x <- compadre$matrixClass[[100]]
# which(grepl("seed",x))
# rm(x)

CompadreSPP <- compadre$metadata[compadre$metadata$AnnualPeriodicity == 1 &
                                   compadre$metadata$MatrixTreatment == "Unmanipulated" &
                                   compadre$metadata$MatrixSplit == "Divided" &
                                   compadre$metadata$MatrixDimension,]

SPP_compadre <- names(table(CompadreSPP$SpeciesAccepted)[table(CompadreSPP$SpeciesAccepted)>4])
SPP_compadretable <- compadre$metadata[compadre$metadata$SpeciesAccepted %in% SPP_compadre,]

# find seed stages; 
taxaSeedStages <- compadre$matrixClass[compadre$metadata$SpeciesAccepted %in% SPP_compadre]

# test
# x <- taxaSeedStages[[1]]
# nr(x)
Nx_taxaCompadre <- lapply(taxaSeedStages, function(x){
  nx <- rep(1,nrow(x))
  nx[x[,1] == "prop"] <- NA
  nx
})

Mx_all_taxaCompadre <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% SPP_compadre], function(x) x$matA)
TMx_all_taxaCompadre <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% SPP_compadre], function(x) x$matU)

# Generation time for each species, lapply over each species
generationtimeCOMPADRE <- do.call(rbind,lapply(SPP_compadre[c(1:3,5:150)], function(x){
  Mx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% x], function(x) x$matA)
  TMx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% x], function(x) x$matU)

  generationtimeCOMPADRE <- mapply(function(Mx,Tx){
    Fm <- Mx - Tx
      popbio::generation.time(Mx, # r = Fm)
      r = c(unique(which(Fm > 0, arr.ind = TRUE)[,1])),
      c = c(unique(which(Fm > 0, arr.ind = TRUE)[,2])))
    }, Mx_one_taxa, TMx_one_taxa)
  data.frame(GT = generationtimeCOMPADRE, SPP = x)
  }))
# library(Rmisc)
gensummary <- summarySE(generationtimeCOMPADRE[!is.infinite(generationtimeCOMPADRE$GT),], measurevar = "GT", groupvars = "SPP", na.rm = TRUE)

longslow <- gensummary[gensummary$GT > 5,]

pathstartRandomCompadre <- "C:/Users/DePrengm/Denver Botanic Gardens/Conservation - Demography/SeedHarvestModelling/Seed_Harvest_Modeling/Simulation_output_COMPADRE/"
# error in simulation longslow$SPP[13:39]
sppnow <- "Actaea cordifolia"
testStages <- compadre$matrixClass[compadre$metadata$SpeciesAccepted == sppnow]
testA <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted == sppnow], function(x) x$matA)
testU <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted == sppnow], function(x) x$matU)
 taxaSeedStages <- compadre$matrixClass[compadre$metadata$SpeciesAccepted == sppnow]
  testNx_list <- lapply(taxaSeedStages, function(x){
    nx <- rep(1,nrow(x))
    nx[x[,1] == "prop"] <- NA
    nx
    })
lapply(testNx_list, function(x) length(x))

length(testNx_list)
length(testA)
length(testU)
testStages[[1]]

# mapply(function(s,A,U) , testStages, testA, testU)

test[!sapply(test, function(x) any(is.na(x)))]
rm(test)


lapply(SPP_compadre[!(SPP_compadre %in% c("Antirrhinum subbaeticum", "Agrimonia eupatoria", "Actaea spicata"))][30:194], function(sp){
  Mx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% sp], function(x) x$matA)
  Mx_one_taxa <- Mx_one_taxa[!sapply(Mx_one_taxa, function(x) any(is.na(x)))]

  TMx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% sp], function(x) x$matU)
  TMx_one_taxa <- TMx_one_taxa[!sapply(TMx_one_taxa, function(x) any(is.na(x)))]

  Nx_one_taxa <- Nx_taxaCompadre[compadre$metadata$SpeciesAccepted %in% sp]
  
  SimSimple("COMPADREslow3gen", sppVector = sp, Simlength = 3, 
          Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = TRUE, ps = pathstartRandomCompadre)
  
  SimSimple("COMPADREslow3gen10yrs", sppVector = sp, Simlength = 10, 
            Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)
  
  SimSimple("COMPADREslow3gen100yrs", sppVector = sp, Simlength = 100, 
            Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)

})

# find this list by see for which the largest reproductive to prop is zero?

# Various matrix sizes, need to have ones that match
#test
# sp <- SPP_compadre[[1]] "Alliaria petiolata"
# stages <- 6 3
# rm(stages)
# test
# x <- taxaSeedStages[[1]]
# nr(x)
rm(taxaSeedStages)
rm(Nx_one_taxa)
rm(sp)

lapply(SPP_compadre, function(sp){
  taxaSeedStages <- compadre$matrixClass[compadre$metadata$SpeciesAccepted == sp]
  Nx_one_taxa <- lapply(taxaSeedStages, function(x){
    nx <- rep(1,nrow(x))
    nx[x[,1] == "prop"] <- NA
    nx
    })
  stageNums <- unlist(lapply(Nx_one_taxa, function(x) length(x)))
  # rle(stageNums)$values[rle(stageNums)$lengths >4] The stage numbers with at least 5 matrices of that size
  lapply(rle(stageNums)$values[rle(stageNums)$lengths >4], function(stages){
    Mx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted == sp], function(x) x$matA)[unlist(lapply(Nx_one_taxa, function(x) length(x) == stages))]
    Mx_one_taxa <- Mx_one_taxa[!sapply(Mx_one_taxa, function(x) any(is.na(x)))]
    TMx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted == sp], function(x) x$matU)[unlist(lapply(Nx_one_taxa, function(x) length(x) == stages))]
    TMx_one_taxa <- TMx_one_taxa[!sapply(Mx_one_taxa, function(x) any(is.na(x)))]
    
    SimSimple("COMPADREslow3gen", sppVector = paste(sp,stages,"stages",sep=""), Simlength = 3, 
            Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = TRUE, ps = pathstartRandomCompadre)
    
    SimSimple("COMPADREslow3gen10yrs", sppVector = paste(sp,stages,"stages",sep=""), Simlength = 10, 
              Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)
    
    SimSimple("COMPADREslow3gen100yrs", sppVector = paste(sp,stages,"stages",sep=""),Simlength = 100, 
              Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)
    })
})


# Semelparous species in COMPADRE from Takada and Kawai 2020
lapply(c("Genianella compestris","Rubus praecox","Rubus ursinus","Carlina vulgaris","Carum carvi","Digitalis purpurea","Picris hieracioides","Centaurea jacea","Cirsium dissectum","Cirsium palustre","Cirsium pitcheri","Cirsium scariosum","Cirsium tracyi","Dipsacus fullonum","Saussurea medusa","Tragapogon pratensis"), function(sp){
  Mx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% sp], function(x) x$matA)
  Mx_one_taxa <- Mx_one_taxa[!sapply(Mx_one_taxa, function(x) any(is.na(x)))]

  TMx_one_taxa <- lapply(compadre$mat[compadre$metadata$SpeciesAccepted %in% sp], function(x) x$matU)
  TMx_one_taxa <- TMx_one_taxa[!sapply(TMx_one_taxa, function(x) any(is.na(x)))]

  Nx_one_taxa <- Nx_taxaCompadre[compadre$metadata$SpeciesAccepted %in% sp]
  
  SimSimple("COMPADREslow3gen", sppVector = sp, Simlength = 3, 
          Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = TRUE, ps = pathstartRandomCompadre)
  
  SimSimple("COMPADREslow3gen10yrs", sppVector = sp, Simlength = 10, 
            Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)
  
  SimSimple("COMPADREslow3gen100yrs", sppVector = sp, Simlength = 100, 
            Mxs = Mx_one_taxa,TMxs = TMx_one_taxa, Nxs = Nx_one_taxa, generationspan = FALSE, ps = pathstartRandomCompadre)

})

```



```{r}
# Skip for now, dormant and damaged stages. 
compadre[[2]][compadre[[1]]$SpeciesAccepted == "Actaea elata"]

Nx_all_ACEL <- list(rep(1,11),rep(1,11),rep(1,11),rep(1,11)) # no seedbank, NA - unobserved seedbank
Mx_all_ACEL <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Actaea elata"], function(x) x$matA))
TMx_all_ACEL <- list(lapply(compadre$mat[compadre$metadata$SpeciesAccepted == "Tillandsia brachycaulos"], function(x) x$matU))


SimSimple("AnnualEmpericalNone", sppVector = "SEER", Simlength = 3, Mxs = Mx_all_SEER, TMxs = TMx_all_SEER, Nxs = Nx_all_SEER, generationspan = TRUE, ps = pathstartempirical3gen)
SimSimple("AnnualEmpericalNone", sppVector = "SEER", Simlength = 10, Mxs = Mx_all_SEER, TMxs = TMx_all_SEER, Nxs = Nx_all_SEER, generationspan = FALSE, ps = pathstartempirical)
SimSimple("AnnualEmpericalNone", sppVector = "SEER", Simlength = 100, Mxs = Mx_all_SEER, TMxs = TMx_all_SEER, Nxs = Nx_all_SEER, generationspan = FALSE, ps = pathstart100)

compadre[[2]][compadre[[1]]$SpeciesAccepted == "Centaurea corymbosa"] # might be semelparous but only one matrix
compadre[[2]][compadre[[1]]$SpeciesAccepted == "Furcraea parmentieri"] # none
compadre[[2]][compadre[[1]]$Genus == "Furcraea"] # none


```













